---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
```

1. For each of the datasets below, specify whether it is in a vector or raster
data format. If it is in a vector data format, explain which types of geometries
it contains (e.g., a point or linestring). Explain your reasoning.
  
  a. [NYC Building Footprints](https://uwmadison.box.com/shared/static/qfmrp9srsoq0a7oj0e7xmgu5spojr33e.geojson)
  
  b. [Africa Population 2020]()
  
  c. [Himalayan Glacial Lakes]()
  
  d. [US EV Charging ]()

  e. [Zion Elevation]()
  
1. For each of the commands below, explain (i) whether it can be applied to
vector data, raster data, or both and (ii) discuss a typical use case for the
function.

    a. `read_sf`
    
    b. `raster`
    
    c. `as.data.frame(img, xy = TRUE)
    
    d. `geom_sf`
    
    e. `coord_sf`

1. [Glacial Lakes]. The data at this link contain labels of glacial lakes the
Hindu Kush Himalaya, created during an ecological survey in 2015 by the
International Centre for Integrated Mountain Development.

  a. How many lakes are in this dataset? What are the latitude / longitude
  coordinates of the largest lakes in each Sub-basin?
  
  b. Plot the polygons associated with each of the lakes identified in step (a).
  _Hint: To make the polygons legible, use `wrap_plots` in the `patchwork` package to combine separate plots for each lake._
  
  c. Visualize the lakes with latitude between 28.2 and 28.4 and with longitude
  between 85.8 and 86. Optionally, add a basemap associated with each lake.
  
1. The problem below visualizes landcover and population data from the
`afrilearndata` package.

    a. What are the dimensions (in pixels) of the population dataset? What is
    its coordinate reference system?
    
    b. Convert the raster data to a `data.frame`. Omit rows without any
    population data.
    
    c. Visualize the data. What are some regions with the lowest and highest
    population densities?
 
1. This problem will visualize the temperature data downloaded from [this
site](http://www.bom.gov.au/jsp/awap/temp/index.jsp?colour=colour&time=latest&step=0&map=maxave&period=month&area=nat)
from the Australian Bureau of Meteorology. It presents some new ideas related to
raster masking.

  a. Read in the raster data, directly convert it to a data.frame, and use a
  `geom_raster` layer to visualize raw temperature data.
  
  b. The data includes temperature over the ocean! We really only want to
  visualize the temperatures over land. Using the `mask` function from the
  `raster` package, clip the original raster data to a polygon containing the
  Australia country boundaries. These boundaries are read in the code block
  below,
      
    ```{r}
    library(rnaturalearth)
    australia <- ne_countries(country = "australia")
    # mask(... fill this in...)
    ```
  
  c. Regenerate the plot from (a), but using the masked raster from (b). Ensure
  that the masked NA values are left out of the visualization. Bonus: Use a
  binned color scale (e.g., `scale_fill_viridis_b`) to make the temperature
  variations even clearer.

1. [Interactive NYC Trees] This problem walks through the creation of an
interactive map of the [NYC Trees Dataset](https://uwmadison.box.com/shared/static/t1mk6i4u5ks5bjxaw2c7soe2z8i75m2o.csv).

  a. Filter down to only those trees whose diameter at breast height
    (`tree_dbh`) is larger than 10. This will help reduce overplotting.
    
  b. Use `leaflet` and the `addCircles` command to create a map with the trees
    kept in part (a). Have the size of the circles reflect the `tree_dbh` value.
    
  c. Add a basemap that highlights features of the city without being too
    busy. For example, you can use
    `addProviderTiles(providers$CartoDB.Positron)`.
    
  d. Color each tree's circle according to its species group. __Hint: Create a mapping from species groups to colors using the `colorFactor` function in Leaflet. For example, `pal <- colorFactor(brewer, filtered_trees$species_group)` creates an objet `pal` that can be called in the `color` argument to `addCircles`.__
  
```{r}
trees <- read_csv("https://uwmadison.box.com/shared/static/t1mk6i4u5ks5bjxaw2c7soe2z8i75m2o.csv")
roads <- read_sf("https://uwmadison.box.com/shared/static/28y5003s1d0w9nqjnk9xme2n86xazuuj.geojson")
```


```{r}
large_trees <- trees %>%
  filter(tree_dbh > 10)

library(leaflet)
brewer <- RColorBrewer::brewer.pal(8, "Set3")
factpal <- colorFactor(brewer, large_trees$species_group)

leaflet(large_trees) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircles(~ longitude, ~ latitude, radius = ~ tree_dbh, col = ~factpal(species_group))
```

```{r}
ggplot(trees) +
  geom_sf(data = roads, fill = "#0C0C0C") +
  geom_point(
    aes(longitude, latitude, col = species_group), 
    size = 0.6, alpha = 0.6
  ) +
  theme_void()
```
 

```{r}
pop <- raster("data/afripop2020.tif") %>%
  as.data.frame(xy = TRUE) %>%
  filter(!is.na(afripop2020))
ggplot(pop) +
  geom_raster(aes(x, y, fill = log(1 + afripop2020))) +
  coord_sf() +
  theme(
    panel.background = element_rect(fill = "transparent"),
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    axis.title = element_blank(),
    legend.position = "bottom"
    )
```
  
  
```{r}
lakes <- read_sf("data/glacial_lakes/GL_3basins_2015.shp")
lakes %>%
  group_by(Sub_Basin) %>%
  filter(Area == max(Area)) %>%
  select(Latitude, Longitude, Sub_Basin, Area) %>%
  arrange(-Area)
```
```{r}
lakes %>%
  filter(
    Latitude < 28.4, Latitude > 28.2,
    Longitude < 86, Longitude > 85.8
    ) %>%
  ggplot() +
    geom_sf(fill = "black", size = 0) +
    theme_void()
```
```{r}
basemap <- get_stamenmap(bbox = c(85.8, 28.2, 86, 28.4))
lakes_zoom <- lakes %>%
  filter(
    Latitude < 28.4, Latitude > 28.2,
    Longitude < 86, Longitude > 85.8
    ) %>%
  rename(lon = Longitude, lat = Latitude)

ggmap(basemap) +
  geom_sf(data = lakes_zoom, fill = "black", size = 0) +
  theme_void()
```
    
```{r, fig.height = 10, fig.width = 10}
top_lakes <- lakes %>%
  group_by(Sub_Basin) %>%
  filter(Area == max(Area)) %>%
  arrange(-Area)

library(patchwork)
p <- list()
for (i in seq_along(top_lakes)) {
  p[[i]] <- ggplot(top_lakes[i, ]) +
    geom_sf(col = "white", fill = "black") +
    theme_void() +
    ggtitle(top_lakes[i, "Sub_Basin"])
}

wrap_plots(p)
```
  
  c. 

```{r}
library(sf)
ev <- read_sf("data/ev/Alternative_Fueling_Stations.shp")
```

```{r}
write_sf(ev[ev$STATE == "WI", ], dsn = "~/Desktop/ev.geojson")
```

```{r}
library(rnaturalearth)
library(scico)

temp <- raster("~/Downloads/latest.grid")
australia <- ne_countries(country = "australia")
temp <- mask(temp, australia) %>%
  as.data.frame(xy = TRUE) %>%
  filter(!is.na(latest))

ggplot(temp) +
  geom_raster(aes(x, y, fill = latest)) +
  scale_fill_viridis_b(option = "magma") +
  theme_void()
```

```{r}
rainfall <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-07/rainfall.csv') %>%
  group_by(lat, long) %>%
  summarise(max = max(rainfall, na.rm = TRUE))

ggplot(rainfall) +
  geom_tile(aes(long, lat, fill = max))

```

```{r}
leaflet(large_trees) %>%
  addProviderTiles(providers$Esri.WorldPhysical) %>%
  addPolylines(data = roads, color = "#0C0C0C", weight = 0.5) %>%
  addCircles(~ longitude, ~ latitude, radius = ~ tree_dbh, col = ~factpal(species_group))

```

