---
author: "Kris Sankaran"
title: "Week10F"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, dpi = 200, fig.align = "center", out.width = 650, fig.height = 3, fig.width = 9)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Uniform Manifold Approximation and Projection

.pull-left[
  April 8, 2022
]
 
---

### Announcements

* Portfolio 3 - April 10
* Project Milestone 3 - April 10
* Practice Midterm 2 is posted
* Midterm 2 is April 18

---

### Today

By the end of the class, you should be able to...

   * Using UMAP, "map" the rows of a high-dimensional, numerical dataset
   * Interpret a UMAP display

---

### Exercise Review

This problem will study a dataset from MIT's [living wage
calculator](https://livingwage.mit.edu/). Each row in this
[dataset](https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week10/exercises/living_wages.csv)
gives living wages for state capitals within the continental US. Since living
wages depend on a personal situation (e.g., number of children), this dataset is
high-dimensional. The interpretation of each column is 
[here](https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week10/exercises/living_wages_lookup.csv).

```{r}
living_wages <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week10/exercises/living_wages.csv")
lookup <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week10/exercises/living_wages_lookup.csv")

head(living_wages, 4)
```

---

a. Define a tidymodels `recipe` that normalizes all city features and
specifies that PCA should be performed.

```{r}
library(tidymodels)

pca_rec <- recipe(~., data = living_wages) %>%
  update_role(state_territory, city, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors())
pca_prep <- prep(pca_rec)
```

---
    
b. Visualize the top 4 principal components. Based on the interpretation of each
columns, what seems to distinguish capital cities with low vs. high values of
PC2?

.pull-left[
```{r, echo = FALSE, fig.height = 8.5}
components <- tidy(pca_prep, 2)
components %>%
  filter(component %in% str_c("PC", 1:4)) %>%
  ggplot() +
  geom_col(aes(value, terms)) +
  facet_grid(component ~ .) +
  theme(axis.text.y = element_text(size = 10))
```
]

.pull-right[
```{r}
lookup
```
]

---
```{r}
renderLeaflet(a)
```

b. Visualize the top 4 principal components. Based on the interpretation of each
columns, what seems to distinguish capital cities with low vs. high values of
PC2?

```{r, eval = FALSE}
components <- tidy(pca_prep, 2)
components %>%
  filter(component %in% str_c("PC", 1:4)) %>%
  ggplot() +
  geom_col(aes(value, terms)) +
  facet_grid(component ~ .) +
  theme(axis.text.y = element_text(size = 10))
```

---
c. Visualize the scores of each capital city with respect to the first two
principal components. Make sure to annotate cities with their names. Pick a city
on the graph and interpret its relative wage profile based on the principal
components.
    
```{r, echo = FALSE, fig.width = 10, fig.height = 5}
scores <- juice(pca_prep)
ggplot(scores) +
  geom_hline(yintercept = 0, size = 2, col = "#5d5d5d") +
  geom_vline(xintercept = 0, size = 2, col = "#5d5d5d") +
  geom_text(aes(PC1, PC2, label = city), size = 5)
```

---
    
c. Visualize the scores of each capital city with respect to the first two
principal components. Make sure to annotate cities with their names. Pick a city
on the graph and interpret its relative wage profile based on the principal
components.

```{r, eval = FALSE, fig.width = 10, fig.height = 5}
scores <- juice(pca_prep)
ggplot(scores) +
  geom_hline(yintercept = 0, size = 2, col = "#5d5d5d") +
  geom_vline(xintercept = 0, size = 2, col = "#5d5d5d") +
  geom_text(aes(PC1, PC2, label = city), size = 5)
```

---

## Live Coding Example

We will work through Exercise 14 in Module 3.

---

### Exercise

* Exercise 10.2 [Phoneme Identification, b-c] on Canvas
* Can discuss, but submit individually