---
author: "Kris Sankaran"
title: "Week10M"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, dpi = 200, fig.align = "center", out.width = 650, fig.height = 3, fig.width = 9)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Principal Components

.pull-left[
  April 4, 2022
]
 
---

### Announcements

* No lecture on Wednesday April 6
* Portfolio 3 - April 10
* Project Milestone 3 - April 10
* Practice Midterm 2 is posted
* Midterm 2 is April 18 (2 weeks)

---

### Today

By the end of the class, you should be able to...

   * Using PCA, "map" the rows of a high-dimensional, numerical dataset
   * Design and interpret a visualization of PCA component directions

---

### Exercise Review

[Polio incidence] In this problem, we will use a heatmap to visualize a large
collection of time series. The
[data](https://uwmadison.box.com/s/nm7yku4y9q7ylvz5kbxya3ouj2njd0x6), prepared
by the [Tycho Project](https://www.tycho.pitt.edu/), contain weekly counts of
new polio cases in the United States, starting as early as 1912 (for some
states).

```{r}
polio <- read_csv("https://uwmadison.box.com/shared/static/nm7yku4y9q7ylvz5kbxya3ouj2njd0x6.csv")
head(polio)
```

---
    
a. Pivot the raw dataset so that states appear along rows and weeks appear along
columns. Fill weeks that don't have any cases with 0's.

```{r}
polio_matrix <- polio %>%
  pivot_wider(state, names_from = period_start_date, values_from = cases, values_fill = 0) %>%
  column_to_rownames("state")
polio_matrix[1:10, 1:4]
```

---

b. Use the `superheat` package to make a heatmap of the data from (a). Have the
color of each tile represent `log(1 + cases)`, rather than the raw counts.
Reorder the states using a hierarchical clustering by setting `pretty.order.rows
= TRUE`.

```{r, fig.width = 12, fig.height = 6, echo = FALSE, out.width = 890}
library(superheat)
cols <- c('#f5eff7','#bdc9e1','#67a9cf','#1c9099','#016c59')
superheat(
  log(1 + polio_matrix),
  left.label.text.size = 3, 
  heat.pal = cols, 
  heat.lim = c(0, 5),
  pretty.order.rows = TRUE
)
```

---

b. Use the `superheat` package to make a heatmap of the data from (a). Have the
color of each tile represent `log(1 + cases)`, rather than the raw counts.
Reorder the states using a hierarchical clustering by setting `pretty.order.rows
= TRUE`.

```{r, fig.width = 12, fig.height = 6, eval = FALSE}
library(superheat)
cols <- c('#f5eff7','#bdc9e1','#67a9cf','#1c9099','#016c59')
superheat(
  log(1 + polio_matrix),
  left.label.text.size = 3, 
  heat.pal = cols, 
  heat.lim = c(0, 5),
  pretty.order.rows = TRUE
)
```

---

c. Supplement the view from part (b) with a barchart showing the US total
incidence during every given week. Interpret the resulting visualization.
*Hint: use the `yt` argument of superheat.*

```{r, fig.width = 12, fig.height = 6, echo = FALSE, out.width = 950}
superheat(
  log(1 + polio_matrix),
  left.label.text.size = 2.4, 
  heat.pal = cols, 
  pretty.order.rows = TRUE,
  yt = colSums(polio_matrix),
  yt.bar.col = "#0c0c0c",
  yt.plot.type = "bar",
  yt.axis.name.size = 0,
  row.title.size = 0,
  legend = FALSE
)
```

---

c. Supplement the view from part (b) with a barchart showing the US total
incidence during every given week. Interpret the resulting visualization.
*Hint: use the `yt` argument of superheat.*

```{r, fig.width = 12, fig.height = 6, eval = FALSE}
superheat(
  log(1 + polio_matrix),
  left.label.text.size = 2.4, 
  heat.pal = cols, 
  pretty.order.rows = TRUE,
  yt = colSums(polio_matrix),
  yt.bar.col = "#0c0c0c",
  yt.plot.type = "bar",
  yt.axis.name.size = 0,
  row.title.size = 0,
  legend = FALSE
)
```

---

## Live Coding Example

We will work through Exercise 2 in Module 3.

---

### Exercise

* Exercise 10.1 [Living Wages] on Canvas
* Can discuss, but submit individually