---
author: ""
title: "Week 12W"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

1. [A Bikesharing Model] In this and the next problem, we will visualize models
fitted to predict bikesharing demand in a subset of the Capitol Bikesharing
[dataset](https://uwmadison.box.com/shared/static/aa91qdqehagag8wg8mqsm4z5b4g2hu0x.csv).
We will see what types of features are learned by different types of models,
whether there are any interactions between features, and whether linear and
nonlinear approaches are substantively different.

a. The code below fits a gradient boosting machine (GBM) to predict bikesharing
demand (count) using all features available, except day of the year. Visualize
the Ceteris Paribus profiles for the temperature and humidity variables, and
provide a brief interpretation.
```{r}
bike
```
      
```{r, echo = TRUE, include = TRUE}
library(caret)
bike <- read_csv("https://uwmadison.box.com/shared/static/aa91qdqehagag8wg8mqsm4z5b4g2hu0x.csv")
x <- dplyr::select(bike, -count, -dteday)
hyper <- data.frame(n.trees = 100, interaction.depth = 4, shrinkage = 0.1, n.minobsinnode = 10)
fit <- train(x = x, y = bike$count, method = "gbm", tuneGrid = hyper, verbose = FALSE)
```
```{r}
library(DALEX)
explanation <- explain(fit, data = x, y = bike$count)
profiles <- model_profile(explanation)
profiles
```

```{r}
plot(profiles, variables = c("temp", "hum"), geom = "profiles") +
  th
```

b. Has the GBM learned an interaction effect between the hour of day (`hr`) and
weekend (`weekend`) features? Briefly justify your answer using a grouped CP
plot.

```{r}
profiles <- model_profile(explanation, groups = "weekend")
plot(profiles, variables = "hr", geom = "profiles") +
  th

```




