---
author: ""
title: "Week 12F"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

[Contrastive Profiles for Bikesharing] This problem continues the exploration in
the previous one. Here, we study whether the choice of GBM was critical, or
whether any other model would have learned the same relationship between the
predictors and bikesharing demand.

a. The code below fits a lasso model to the same input and output dataset.
Provide contrastive partial dependence profiles between the lasso and the GBM
from the previous problem, focusing on the hour (`hr`), humidity (`hum`), and
temperature (`temp`) features. Comment on the result.

```{r, echo = TRUE, include = TRUE}
library(tidyverse)
library(caret)

bike <- read_csv("https://uwmadison.box.com/shared/static/aa91qdqehagag8wg8mqsm4z5b4g2hu0x.csv")
x <- select(bike, -count, -dteday)

fits <- list()
hyper <- data.frame(n.trees = 100, interaction.depth = 4, shrinkage = 0.1, n.minobsinnode = 10)
fits[["gbm"]] <- train(x = x, y = bike$count, method = "gbm", tuneGrid = hyper, verbose = FALSE)

hyper <- data.frame(lambda = 0.001, alpha = 0)
fits[["lm"]] <- train(x = x, y = bike$count, method = "glmnet", tuneGrid = hyper)
```
 
 
 
 
b. Repeat part (a), but comparing the GBM with the CART model fitted below.
Comment on the result.
  
```{r, echo = TRUE, include = TRUE}
hyper <- data.frame(cp = 0.01)
fits[["cart"]] <- train(x = x, y = bike$count, method = "rpart", tuneGrid = hyper)
```