---
title: "STAT 479: MIDTERM 2"
geometry: margin=2cm
output: 
  pdf_document:
    keep_tex: true
---
```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(raster)
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = TRUE)
```
\rule[-1mm]{19.3cm}{1mm}
Name: \rule[0pt]{7cm}{.5pt}\vspace{3mm}

* This exam lasts from 1:20 - 2:10pm on April 18, 2022. There are 6 questions.
* This exam is closed notes and closed computer.
* You may use a 1-page cheat sheet (8.5 x 11in or A4 size). You may use both
sides, but the cheat sheet must be handwritten.
* If you need extra space, you may write on the back of the page. Please
indicate somewhere that your answer continues.
* The instructors will only be able to answer clarifying questions during the
exam. They will be sitting at the back of the room.

| Question | Q1 | Q2 | Q3 | Q4 | Q5 | Q6 | Total |
| ---- | -- | -- | -- | -- | -- | -- | -- |
| Score |  |  |  |  |  |  |  |  | |
| Possible |  4 | 4 | 5 | 6 | 5 | 6 | 30  |

\rule[1mm]{19.3cm}{1mm}

### Q1 (4 points)

Circle TRUE or FALSE for all statements below about time series and geospatial
visualization.

  * TRUE or FALSE: The `geom_sf` layer can be used to visualize collection of
  spatial polygons of class `sf`.
  * TRUE or FALSE: The `geom_raster` layer can be used to visualize a
  multi-channel raster as an RGB image.
  * TRUE or FALSE: A gradually decreasing slope in an ACF visualization
  indicates a time series with strong seasonal structure.
  * TRUE or FALSE: If `x` is a data.frame with columns `date`, `city`, and
  `temperature`, then command `as_tsibble(x, key = city, index = date)` will
  correctly create a `tsibble` object containing the time series collection,
  assuming there are no gaps in time.
  * TRUE or FALSE: A per-pixel grid of temperature data is an example of a
  spatial vector dataset.

### Q2 (4 points)

Circle TRUE or FALSE for all statements below about clustering and topic models.
    
  * TRUE or FALSE: If all samples in a cluster have large silhouette statistics, then that
  cluster is poorly-defined.
  * TRUE or FALSE: Consider the `survey` dataset below,
        
    ```
      Country     `WBL Report Year` Question answer
      <chr>                   <dbl> <fct>     <dbl>
    1 Afghanistan              1971 Q1            0
    2 Afghanistan              1971 Q2            1
    3 Afghanistan              1971 Q3            0
    4 Afghanistan              1971 Q4            1
    ```
  
    To use $K$-means to cluster country $\times$ year combinations by their pattern of question responses, we would need to first pivot it using the command,
      
    ```{r, eval = FALSE}
    survey %>%
    	pivot_wider(names_from = "Country", values_from = "answer")
    ```	
      
  * TRUE or FALSE: Suppose a data.frame has a column, `reviews`, containing raw
  text from across a collection of reviews (one review per row). Then, the
  `cast_dtm` function can be used to split the raw text so that one word appears
  on each row.
  * TRUE or FALSE: A Structure plot visualizes the `n_samples` $\times$
  `K`-dimensional matrix of mixed memberships fitted by a topic model.
  * TRUE or FALSE: When cutting a hierarchical clustering tree at two heights,
  the clusters found at the height closer to the leaves will always be nested
  within the clusters found at the height closer to the root.
  
### Q3 (5 points)

Imagine that you have been tasked with training models of glacial change in the
Himalaya mountain range. You decide to first generate exploratory visualizations
of the relevant spatial data.

a. [2 points] The block below reads in an `sf` object with one row per glacier.
Provide code for visualizing these glaciers, faceted by the `Glaciers` category.
An example result is given in Figure \ref{fig:glaciers_sf}.
    
```{r}
library(sf)
glaciers <- read_sf("https://uwmadison.box.com/shared/static/k6htkjqj0w48g8oixfwq0zaobsny04jh.geojson") %>%
  filter(Longitude < 86.91, Longitude > 86.1, Latitude < 27.81, Latitude > 27.7)

glaciers[1:2, 1:5]
```
  
```{r, fig.cap = "An example result for Q3a. \\label{fig:glaciers_sf}", fig.align = "center", out.width = "0.5\\textwidth", echo = FALSE}
ggplot(glaciers) +
  geom_sf() +
  facet_wrap(~ Glaciers)
```

$$\\[1in]$$

b. [1 point] The block below reads in a raster image for the same region. The
15th channel of this dataset contains information about the slope at each pixel
in the image. Provide code that could be used to visualize the slope at each
location in the dataset.

```{r}
image <- brick("https://uwmadison.box.com/shared/static/2z3apyg4t7ct5qd4mcwh9rpr63t02jql.tif")
```  

$$\\[1in]$$

c. [2 points] Suggest an approach to layering the glacier boundaries from (a)
onto the slope map from (b). An example result is given in Figure
\ref{fig:glaciersoverlay}.

```{r, fig.cap = "\\label{fig:glaciersoverlay} An example result for Q3c.", fig.align = "center", out.width = "0.5\\textwidth", echo = FALSE}
library(RStoolbox)
ggRGB(image, r = 15, g = 15, b = 15, alpha = 0.3) +
  geom_sf(data = glaciers, col = "red") +
  facet_wrap(~ Glaciers)
```

$$\\[1in]$$

### Q4 (6 points)

In this problem, we will see how a heatmap can visualize user ratings across a
collection of movies.

```{r}
movies_mat <- read_csv("https://uwmadison.box.com/shared/static/wj1ln9xtigaoubbxow86y2gqmqcsu2jk.csv") %>%
  column_to_rownames(var = "title")
movies_mat[1:2, 1:10]
```

a. [2 points] Provide code to generate the heatmap shown in Figure \ref{fig:movies}.

```{r, echo = FALSE, fig.height = 10, fig.width = 18, out.width = "0.6\\textwidth", fig.cap = "\\label{fig:movies} A heatmap of movie ratings for Q4a.", fig.align = "center"}
library(superheat)
superheat(movies_mat, pretty.order.rows = TRUE, pretty.order.cols = TRUE)
```

$$\\[2in]$$

b. [2 points] In this visualization, how do you interpret sets of similar rows?
Similar columns?

$$\\[1in]$$

c. [2 points] Propose an alternative compound plot that also allows readers to
easily find average movie ratings. Draw a simple sketch of what the output would
look like and describe any necessary modifications to the code in part (a).

$$\\[1in]$$

### Q5 (5 points)

The code below reads in nodes and edges for a graph describing the similarity of
512 HIV samples collected across 32 countries. Two samples are linked if they
have similar genetic sequences.

```{r, echo = TRUE, warning = FALSE, message = FALSE}
library(tidygraph)
nodes <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week11/exercises/hiv_nodes.csv")
edges <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week11/exercises/hiv_edges.csv")
G <- tbl_graph(nodes, edges)
G
```

a. [2 points] Provide code to generate a node-link visualization similar to that
in Figure \ref{fig:hiv_graph}. Make at least one style customization that
improves the visualization. Justify your choice.

```{r, echo = FALSE, fig.width = 10, fig.height = 10, out.width = "0.5\\textwidth", fig.cap = "\\label{fig:hiv_graph} A network between HIV sequences, labeled by the country from which they originated.", fig.align = "center"}
library(ggraph)
ggraph(G) +
  geom_edge_link(colour = "#d3d3d3", alpha = 0.9, width = 2) +
  geom_node_text(aes(label = name)) +
  theme_void()
```

$$\\[1.5in]$$
  
b. [1 point] The figure below shows the node-link visualization for a simple
graph. Sketch its exact adjacency matrix.

```{r, echo = FALSE, out.width = "0.3\\textwidth", fig.width = 5, fig.height = 5, fig.align = "left"}
set.seed(123)
G <- as_tbl_graph(sample_pa(5)) %>%
  mutate(node = 1:5)
ggraph(G, layout = "kk") +
  geom_edge_link() +
  geom_node_label(aes(label = node)) +
  theme_void()
```

$$\\[1.5in]$$

c. [2 points] The figure below shows the node-link visualization for a
more complex graph. How do you expect its corresponding adjacency matrix
visualization to look? Provide a rough sketch and a one-sentence justification.

```{r, echo = FALSE, out.width = "0.3\\textwidth", fig.width = 10, fig.height = 10, fig.align = "left"}
G <- as_tbl_graph(sample_islands(3, 40, .12, 2))
ggraph(G) +
  geom_edge_link() +
  geom_node_point() +
  coord_fixed() +
  theme_void()
```

$$\\[1.5in]$$

### Q6 (6 points)

This problem asks you to explain, in your own words, a few conceptual and
practical aspects for **one of the following** dimensionality reduction
techniques,

  * Principal Components Analysis (PCA)
  * Uniform Manifold Approximation and Projection (UMAP)
  * Latent Dirichlet Allocation (LDA)
  
a. [3 points] Outline how you would implement this method in R and extract
output needed for downstream visualizations. Make sure to highlight the the most
important steps and functions and briefly explain any code you write down.

$$\\[4in]$$

b. [3 points] Describe the outputs from (a). How is each part interpreted? Be as
specific as possible.

$$\\[4in]$$