---
title: "STAT 479: PRACTICE MIDTERM 1"
author: "STAT 479"
date: "February 21, 2022"
output:
  pdf_document:
    keep_tex: true
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
opts_chunk$set(cache = TRUE, message = FALSE, warning = FALSE, echo = TRUE, include = TRUE)
options(width = 200)
```

* This exam lasts from 1:20 - 2:10pm on February 28, 2022. There are 8
questions.
* This exam is closed note and closed computer.
* You may use a 1-page cheat sheet (8.5 x 11in or A4 size). You may use both
sides, but the cheat sheet must be handwritten.
* If you need extra space, you may write on the back of the page. Please
indicate somewhere that your answer continues.
* The instructors will only be able to answer clarifying questions during the
exam. They will be sitting at the back of the room.

| Question | Q1 | Q2 | Q3 | Q4 | Q5 | Q6 | Q7 | Q8 | Total |
| ---- | -- | -- | -- | -- | -- | -- | -- | -- | -- |
| Score |  |  |  |  |  |  |  |  | |
| Possible | 2 | 2 | 3 | 3 | 5 | 3 | 6 | 6 | 30 |


### Q1 [2 points]

Circle all the statements below that are true about small-multiple and compound
figures. Assume that the `ggplot2`, `ggridges`, and `patchwork` packages have
been loaded.

  a. Each panel in a plot using `+ facet_grid(x ~ y)` will show a
  different subset of the data.
  
  b. The number of vertical panels in a plot using `facet_wrap(~ x)` depends
  on the number of levels of `x`.
  
  c. If `p1`, `p2`, and `p3` are each `ggplot2` objects, then using `p1 + p2 + p3` 
  will combine those plots horizontally.
  
  d. The ridge heights across `x` in a plot using `+ geom_ridge_density(aes(x, y))` encodes the variable `y`.

### Q2 [2 points]

1. On the line, enter the letter that best matches the following definitions. A
term may be used more than once or not at all.

    * __ The use of the same visual encoding to several partitions of the
    dataset, shown together across adjacent panels.
    * __ A detail view that appears when one or more items are selected in the main
    view.
    * __ The use of different visual encodings to represent the same data, shown
    together across adjacent panels.
    * __ A practice where items interactively selected in one overview are
    highlighted in all other views.

    A. Ridgeline plot
    B. Multiform display
    C. Details-on-demand
    D. Small multiples
    E. Linked brushing
    F. Tidy data
    
\newpage

### Q3 [3 points]

Discuss two common pitfalls in compound figures. What can be done to avoid them?

$$\\[3in]$$
    
### Q4 [3 points] 
    
Sketch the code needed to transform `medals`,

```{r, echo = FALSE}
medals <- data.frame(
  "country" = c("USA", "Canada"),
  "gold" = c(10, 20),
  "silver" = c(7, 26)
)
medals
```

into the form given below,

```{r, echo = FALSE}
library(tidyverse)
medals %>%
  pivot_longer(gold:silver, "medal", "count")
```


\newpage

### Q5 [5 points]

The questions below refer to the two figures below.

```{r, out.width = "50%", fig.show = "hold", echo = FALSE}
library(EBImage)
display(readImage("https://krisrs1128.github.io/stat479/posts/2021-01-12-week1-2/week1-2_files/figure-html5/unnamed-chunk-8-1.png"))
display(readImage("https://krisrs1128.github.io/stat479/posts/2021-01-12-week1-2/week1-2_files/figure-html5/unnamed-chunk-16-1.png"))
```

a. [2 points] Describe three distinct customizations used to transform the plot
on the left into the plot on the right.
$$\\[2in]$$
    

b. [3 points] Provide the `ggplot2` commands that were used for each
transformation. Be as specific as possible.
$$\\[3in]$$
    
\newpage

### Q6 [3 points]
    
This problem asks you to compare and contrast the `geom_ridge_density()` and
`geom_rideline()` layers available in the `ggridges` package.

a. [3 points] Describe the differences in the visual encodings provided by these
two layers. What kind of data do they assume, and how are they represented
graphically?

$$\\[2in]$$


b. [2 points] For each type of layer, provide one situation where one could be
applied, but not the other.

$$\\[2.5in]$$

### Q7 [6 points]

Provide the reactivity graph associated with the shiny app below, which prints
the summary and first few rows of one of three datasets, depending on the user's
selection.
    
```{r}
library(shiny)
ui <- fluidPage(
    textInput("caption", "Caption:", value = "Data Summary"),
    selectInput("dataset", "Dataset:", choices = c("rock", "pressure", "cars")),
    numericInput("obs", "Number of observations:", value = 10),
    verbatimTextOutput("summary"),
    dataTableOutput("view")
)

server <- function(input, output) {
  datasetInput <- reactive({
    if (input$dataset == "rock") {
      return (rock)
    } else if (input$dataset == "pressure") {
      return (pressure)
    }
    cars
  })

  output$caption <- renderText({ input$caption })
  output$summary <- renderPrint({
    dataset <- datasetInput()
    summary(dataset)
  })

  output$view <- renderDataTable({ head(datasetInput(), n = input$obs)})
}
```

\newpage

### Q8 [6 points]

Consider the dataset below, which contains the average `Asymmetry` across
several groups of birds. Note that only the first few rows are shown,
    
```{r}
  birds <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/exercises/data/birds.csv") %>%
  filter(!is.na(Order))
  
  birds
```
    
a. [3 points] Provide code to convert the `Order` variable to lower case and
to compute the mean `Asymmetry` within each `Order`.
    
$$\\[2in]$$
    
```{r, echo = FALSE}
birds <- birds %>%
  group_by(Order) %>%
  summarise(mean_asymmetry = mean(Asymmetry)) %>%
  mutate(Order = tolower(Order))
```


b. [3 points] Provide code needed to generate the plot below. Be as specific as
possible.
    
```{r, echo = FALSE, fig.height = 4, fig.width = 8}
birds %>%
  ggplot() +
  geom_bar(aes(mean_asymmetry, reorder(Order, mean_asymmetry)), width = 1, stat = "identity") +
  labs(x = "Mean Asymmetry", y = "Order") +
  scale_x_continuous(expand = c(0, 0, 0.1, 0)) +
  theme(axis.ticks = element_blank(), panel.grid = element_blank())
```

$$\\[3in]$$