---
author: "Kris Sankaran"
title: "week5M"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(webshot)
library(tidyverse)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, dpi = 200, fig.align = "center", out.width = 650)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Introduction to Graphical Queries

.pull-left[
February 21, 2022
]
 
---

### Announcements

* Practice midterm posted
  - Exam in-class next Monday (February 28)
* Portfolio peer reviews description posted (February 27)
* Project Milestone 2 is posted, and the deadline is extended (March 13)
  
---

### Today

* Exercise review
  - Different types of plot outputs
  - Example reactivity graphs
* Setup a click-based graphical query
* Explaining use cases for dynamic queries
* Review for the midterm exam

---

## Exercise 4.3 Discussion

* Many creative linked outputs
* Effective code externalization 
  - Clear boundaries between component relations and reaction logic
  - For most submissions, `ui` and `server` was shorter than 30 lines
* Overall, very nice work!

---

### Examples

* Complementing with data.frame or summaries
* Complementing with histograms or scatterplots
  - Can you imagine dynamically linked versions?
* Allowing selection by fire or counties
  - Implements details-on-demand

---

### Debugging Shiny Apps

* `Cmd/Ctrl + Shift + Enter` to quickly relaunch after small changes
* `browser()` to interactively evaluate
  - `n` for next line, `c` to next `browser()`, and `Q` to exit
* `message` and `glue` for checking order of execution
* Prune away until we have a minimal failing app

---


```{r, eval = FALSE}
library(shiny)

ui <- fluidPage(
  sliderInput("x", "x", value = 1, min = 0, max = 10),
  sliderInput("y", "y", value = 2, min = 0, max = 10),
  sliderInput("z", "z", value = 3, min = 0, max = 10),
  textOutput("total")
)
server <- function(input, output, session) {
  observeEvent(input$x, {
    updateSliderInput(session, "y", value = input$x * 2)
  })
  
  total <- reactive({
    browser()
    tmp <- input$x + input$y
    tmp + input$z
  })
  output$total <- renderText(total())
}

shinyApp(ui, server)
```


---

```{r, eval = FALSE}
library(shiny)
library(glue)

ui <- fluidPage(
  sliderInput("x", "x", value = 1, min = 0, max = 10),
  sliderInput("y", "y", value = 2, min = 0, max = 10),
  sliderInput("z", "z", value = 3, min = 0, max = 10),
  textOutput("total")
)
server <- function(input, output, session) {
  observeEvent(input$x, {
    message(glue("Updating y from {input$y} to {input$x * 2}"))
    updateSliderInput(session, "y", value = input$x * 2)
  })
  
  total <- reactive({
    total <- input$x + input$y + input$z
    message(glue("New total is {total}"))
    total
  })
  output$total <- renderText(total())
}

shinyApp(ui, server)
```


---

### Notes review

(go to [link](https://drive.google.com/file/d/19-LUE7kNBCIZLxpXoF69xyEq32PAZBw6/view?usp=sharing))

---

## Exercise

* In this exercise, you will work with your team to prepare two slides
summarizing a topic that may appear on the midterm exam.
* Your slides should cover at least two of the following,
  - Important vocabulary
  - Potential pitfalls or bugs
  - A code example (either from class, or your own)
  - Pointers to relevant readings, notes, or other resources
* Slides [link](https://docs.google.com/presentation/d/12quG_sPmG3qQjLLH2UJ4aP2CcWaWhVf-cw8Xq-1mYPg/edit?usp=sharing)

---

### Exercise

* Exercise 5.1 on Canvas
  - Edit slides [here](https://docs.google.com/presentation/d/12quG_sPmG3qQjLLH2UJ4aP2CcWaWhVf-cw8Xq-1mYPg/edit?usp=sharing)
  - Also upload your two slides to Canvas
* Until 2pm, then share with neighboring team
* Submit solutions as a team

