---
author: "Kris Sankaran"
title: "week5W"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(webshot)
library(tidyverse)
library(shiny)
library(bslib)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = FALSE, dpi = 200, fig.align = "center", out.width = 650, fig.height = 3, fig.width = 9)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Linked Brushing

.pull-left[
  February 23, 2022
]
 
---

### Announcements

* Portfolio peer reviews description posted (February 27)
* Midterm Exam in-class next Monday (February 28)
  
---

### Today

By the end of the class, you should be able to...

* Explain situations where brushed-based graphical queries may be useful
* Implement brushed-based graphical queries
* Apply the `bslib` or shinyThemes libraries to customize app appearance

---

```{r, eval = FALSE}
library(shiny)

ui <- fluidPage(
  titlePanel("Old Faithful Geyser Data"),
  sliderInput("bins", "Number of bins:", min = 1, max = 50, value = 30),
  plotOutput("distPlot")
)

server <- function(input, output) {
  output$distPlot <- renderPlot({
      ggplot(faithful) + geom_histogram(aes(waiting), bins = input$bins)
  })
}

shinyApp(ui, server)
```

---

```{r, echo = FALSE}
library(shiny)

ui <- fluidPage(
  titlePanel("Old Faithful Geyser Data"),
  sliderInput("bins", "Number of bins:", min = 1, max = 50, value = 30),
  plotOutput("distPlot")
)

server <- function(input, output) {
  output$distPlot <- renderPlot({
      ggplot(faithful) + geom_histogram(aes(waiting), bins = input$bins)
  })
}

shinyApp(ui, server)
```

---

```{r, eval = FALSE}
library(shiny)
library(bslib)

ui <- fluidPage(
  titlePanel("Old Faithful Geyser Data"),
  sliderInput("bins", "Number of bins:", min = 1, max = 50, value = 30),
  plotOutput("distPlot"),
  theme = bs_theme(bootswatch = "minty")
)

server <- function(input, output) {
  output$distPlot <- renderPlot({
      ggplot(faithful) + geom_histogram(aes(waiting), bins = input$bins)
  })
}

shinyApp(ui, server)
```

---

```{r, echo = FALSE}
library(shiny)
library(bslib)

ui <- fluidPage(
  titlePanel("Old Faithful Geyser Data"),
  sliderInput("bins", "Number of bins:", min = 1, max = 50, value = 30),
  plotOutput("distPlot"),
  theme = bs_theme(bootswatch = "minty")
)

server <- function(input, output) {
  output$distPlot <- renderPlot({
      ggplot(faithful) + geom_histogram(aes(waiting))
  })
}

shinyApp(ui, server)
```

---

```{r, eval = FALSE}
library(shiny)
library(bslib)

my_theme <- bs_theme(
  bg = "#ffe6dc", fg = "#ff682c",
  base_font = font_google("Shizuru")
)

ui <- fluidPage(
  titlePanel("Old Faithful Geyser Data"),
  sliderInput("bins", "Number of bins:", min = 1, max = 50, value = 30),
  plotOutput("distPlot"),
  theme = my_theme
)

server <- function(input, output) {
  output$distPlot <- renderPlot({
      ggplot(faithful) + geom_histogram(aes(waiting))
  }, bg = "transparent")
}

shinyApp(ui, server)
```

---

```{r, echo = FALSE}
library(shiny)
library(bslib)

my_theme <- bs_theme(
  bg = "#ffe6dc", fg = "#ff682c",
  base_font = font_google("Shizuru")
)

ui <- fluidPage(
  titlePanel("Old Faithful Geyser Data"),
  sliderInput("bins", "Number of bins:", min = 1, max = 50, value = 30),
  plotOutput("distPlot"),
  theme = my_theme
)

server <- function(input, output) {
  output$distPlot <- renderPlot({
      ggplot(faithful) + geom_histogram(aes(waiting))
  }, bg = "transparent")
}

shinyApp(ui, server)
```

---

### Notes review

(go to [link](https://drive.google.com/file/d/1ipjCsQ7zRqS_-2W968bXDJ2ogAN91cMb/view?usp=sharing))

---

### Exercise

* You will develop an application to graphically query Airbnb rental prices in NYC
in 2019.
* Some other potentially interesting queries,
  - Clicking on barchart of number of full apartment, private room, and shared
  room rentals
  - Brushing histograms or scatterplots with numbers of reviews and square footage
  - Clicking or brushing bar charts of number of apartments by neighborhood

---

### Exercise

* Exercise 5.2 on Canvas
  - Hint sets up the code skeleton
* Until 2pm
* Submit solutions individually
