---
author: "Kris Sankaran"
title: "Week12F"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
library(caret)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, dpi = 200, fig.align = "center", out.width = 650, fig.height = 3, fig.width = 9)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Partial Dependence Plots (Part 2)

.pull-left[
  April 20, 2022
]

---

### Announcements

---

### Today

By the end of the class, you should be able to...

  * Implement contrastive CP profiles
  * Discuss the difference between two nonlinear models based on their
  contrastive CP profiles

---

### Exercise Review

[Gender Pay Gap] This problem uses CP profiles to investigate the gender gap
in a Glassdoor dataset of employee salaries. It is helpful to use a model,
because it allows us to control for multiple other factors -- a direct plot of
salary vs. gender could be criticized as not accounting for confounding
variables. The code below trains a gradient boosting machine model on `BasePay`
variable (yearly salary in USD), using all potential predictors in the dataset.
    
```{r, warning = FALSE, message = FALSE}
salary <- read_csv("https://github.com/krisrs1128/stat479_s22/raw/main/_slides/week12/exercises/Glassdoor%20Gender%20Pay%20Gap.csv")
x <- salary %>%
  select(Gender:Seniority) %>%
  mutate(across(where(is.character), as.factor)) # gbm needs chr -> factor
y <- salary$BasePay
fit <- train(x, y, method = "gbm", verbose = FALSE)
```

---

a. Before attempting to explain the model, it is helpful to consider its
accuracy. Make a plot of the truth (`y`) against model predictions (`y_hat <- predict(fit)`) and comment on model performance.

```{r}
data.frame(y = y, y_hat = predict(fit)) %>%
  ggplot() +
    geom_point(aes(y, y_hat), alpha = 0.6, size = 0.5) +
    geom_abline(slope = 1, col = "red", size = 1) +
    coord_fixed()
```


---

b. Compute aggregate CP profiles grouped by gender and comment on the extent
of the gender pay gap. According to the fitted prediction surface, is there
more or less of a pay gap at certain ages or levels of seniority?

```{r}
library(DALEX)
explanation <- explain(fit, data = x, y = y)
```

---

b. Compute aggregate CP profiles grouped by gender and comment on the extent
of the gender pay gap. According to the fitted prediction surface, is there
more or less of a pay gap at certain ages or levels of seniority?

```{r}
profiles <- model_profile(explanation, groups = "Gender")
plot(profiles)
```

There appears to be a gender gap uniformly across all other variables.

---
  
c. Show the analogous display without aggregating (i.e., `geom = "profiles"`). What is the interpretation of each line in this plot? 

```{r}
plot(profiles, geom = "profiles")
```

Each line corresponds to a sample. Its $\left(x, y\right)$ coordinates
correspond to predicted salaries $y$ after intervening on the plotted variable's
$x$.

---


## Live Coding Example

We will work through Exercise  [Contrastive Profiles for Bikesharing] in Module 3.

---

### Exercise

* Exercise 12.2 [Overfitting `sin(x)`] on Canvas
* Can discuss, but submit individually