---
author: "Kris Sankaran"
title: "Week11W"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, dpi = 200, fig.align = "center", out.width = 650, fig.height = 3, fig.width = 9)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Visualizing Topic Models

.pull-left[
  April 13, 2022
]

---

### Announcements

* Practice Midterm 2 is posted
  - I will go over solutions on Friday
* Midterm 2 is April 18
* Portfolio 3 Peer Reviews due April 17
* **Project Peer Reviews due April 20**

---

### Today

By the end of the class, you should be able to...

   * Visualize the memberships and topics fitted by a topic model
   * Interpret the memberships and topics fitted by a topic model

---

### Exercise Review

[Personality Types] This
[dataset](https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week11/exercises/mbti_1.csv?raw=true)
contains a sample of posts from the Personality Cafe forum, together with labels
giving the poster's (self-reported) Meyer-Briggs type (e.g., types starting with
"I" are introverted, those with "E" are extroverted). It is time-consuming to go
through all the posts manually, but to get a quick overview of the main topics
that are discussed, we can use a Latent Dirichlet Allocation (LDA) model. This
problem walks through the steps for (1) preparing a Document-Term matrix for
LDA, (2) fitting the LDA model, and (3) interpreting the estimated topics and
memberships.


---

a. Transform the raw posts into a collection of per-post word counts. Remove
stopwords from across all lexicons in `tidytext::stopwords`.

```{r}
library(tidytext)
mbti <- read_csv("https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week11/exercises/mbti_1.csv?raw=true") %>%
  sample_n(500) %>% # can set sample_n(50) for quick testing
  mutate(document = as.character(row_number()))

mbti
```

---

a. Transform the raw posts into a collection of per-post word counts. Remove
stopwords from across all lexicons in `tidytext::stopwords`.

```{r}
mbti %>%
  unnest_tokens(word, posts) %>%
  filter(!(word %in% stop_words$word))
```

---

a. Transform the raw posts into a collection of per-post word counts. Remove
stopwords from across all lexicons in `tidytext::stopwords`.

```{r}
mbti %>%
  unnest_tokens(word, posts) %>%
  filter(!(word %in% stop_words$word)) %>%
  count(document, word)
```

---

b. Convert the `data.frame` from (a) into a topic models Document-Term
matrix (i.e., an object of class `DocumentTermMatrix`). Fit an LDA model
with 8 topics to the prepared object.

```{r}
mbti_dtm <- mbti %>%
  unnest_tokens(word, posts) %>%
  filter(!(word %in% stop_words$word)) %>%
  count(document, word) %>%
  cast_dtm(word, document, n)

mbti_dtm
```

---

b. Convert the `data.frame` from (a) into a topic models Document-Term
matrix (i.e., an object of class `DocumentTermMatrix`). Fit an LDA model
with 8 topics to the prepared object.

```{r}
library(topicmodels)
fit <- LDA(mbti_dtm, k = 8)
```

---

## Live Coding Example

We will work through Exercise  [Hotel Reviews] in Module 3.

---

### Exercise

* Exercise 11.2 [Midterm Review] on Canvas
* Submit as a group