---
author: ""
title: "Week 13M"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

[Deep Learning as a Mixture of Separating Planes] In this exercise, we will
visualize the predictions and hidden layer activations in a simple, but
nonlinear, classification problem. The goal is to illustrate how nonlinear
decision boundaries can be learned by deeper layers of a neural network, even
though each neuron is a linear combination of the previous layer's units.

a. The block below simulates a toy dataset with predictors `x` and response
`y`. Provide a visualization for the dataset and comment on why a one-layer
deep learning model would not be sufficient to achieve good performance.

```{r}
x <- matrix(rt(N * 2, 12), N, 2)
y <- rowSums(x ^ 2) < 1
flip_ix <- sample(N, N / 20) # 5% unavoidable error
y[flip_ix] <- !y[flip_ix]
simulation <- data.frame(x = x, y = y)
```

b. The block below defines and fits a deep learning model with three hidden
layers. Design and implement a visualization that shows the predicted
probabilities for each sample after the model has converged.

```{r}
library(keras)
model <- keras_model_sequential() %>% 
  layer_dense(units = 20, activation = "relu", input_shape = c(2)) %>% 
  layer_dense(units = 20, activation = "relu", input_shape = c(20)) %>% 
  layer_dense(units = 20, activation = "relu", input_shape = c(20)) %>% 
  layer_dense(2, activation = "softmax")

model %>% 
  compile(optimizer = "adam", loss = "categorical_crossentropy") %>%
  fit(
    x = simulation %>% select(starts_with("x")) %>% as.matrix(),
    y = to_categorical(simulation$y),
    metric = "accuracy",
    epochs = 60,
    verbose = 0
  )
```
  
c. The block below defines a `keras` model that can be used to extract
activations from the first hidden layer of the trained deep learning
model. Design and implement a visualization to characterize how
activations for a single neuron behave, as a function of the input $x$
coordinates. Make the analogous visualization for a deeper layer.

d. Generate a heatmap of activations across all neurons from some layer of
the model. Comment on the structure of the learned features.

