---
author: ""
title: "Week 12W"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(reticulate)
library(tidyverse)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

1. [Histopathology Embeddings] This problem investigates the features learned by
a residual network model trained trained to classify histopathology slides.
Specifically, the script at [this
link](https://colab.research.google.com/drive/1bPFz718F-YE0iUcdkm2kc0Aeg8g6-Neh?usp=sharing)
was used to train a model to images from the Pcam [benchmark
dataset](https://www.kaggle.com/competitions/histopathologic-cancer-detection/data).
Each image is a tissue slide. The class labels describe whether the center $32 \times 32$ patch within the image contains any cancerous cells.

In this problem, we will also practice using the `reticulate` package to read in
numpy arrays produced by the python training script linked above. This language
interoperability makes it possible to use the packages best suited to both
modeling (`pytorch`) and visualization (`ggplot2`).

a. We have hosted a subsample of the training images at [this
link](https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week13/exercises/subset.tar.gz?raw=true).
Their corresponding [labels](https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week13/exercises/y.npy?raw=true) are stored here as numpy arrays. Visualize the raw images corresponding to 10 images from each class. _Hint: To unzip these files from the command line, you can use `tar -zxvf subset.tar.gz`_ 

```{python}
import numpy as np
fnames = np.load("/Users/ksankaran/Downloads/tmp/fnames.npy")
y = np.load("/Users/ksankaran/Downloads/tmp/y.npy")
```

```{r}
library(raster)
library(RStoolbox)

# download the subset.tar.gz and unzip
data_dir <- "/Users/ksankaran/Downloads/subset/"
ix <- which(py$y == 1)[1:10]
ims <- map(py$fnames[ix], ~ brick(file.path(data_dir, .)))
map(ims, ~ ggRGB(.))
```


```{r}
ix <- which(py$y == 0)[1:10]
ims <- map(py$fnames[ix], ~ brick(file.path(data_dir, .)))
map(ims, ~ ggRGB(.))
```
    
b. For the subsample in part (a), we have saved the residual network
features from the final layer (pre-classification) layer. They are available
at [this
link](https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week13/exercises/h.npy?raw=true).
Generate UMAP embeddings for the images based on these features, and shade
in each sample according to its class.

```{python}
h = np.load("/Users/ksankaran/Downloads/tmp/h.npy")
```

```{r}
library(tidymodels)
library(embed)

umap_rec <- data.frame(py$h) %>%
  mutate(
    file = py$fnames,
    y = as.factor(py$y)
  ) %>%
  recipe(~ ., data = .) %>%
  update_role(y, file, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_umap(all_predictors())

umap_prep <- prep(umap_rec)
coords <- juice(umap_prep)
```

```{r, fig.width = 10, fig.height = 10}
p <- ggplot(coords) +
  geom_point(aes(UMAP1, UMAP2, col = y)) +
  coord_fixed()
p
```
    
c. Using the `annotation_raster` function from `ggplot2`, overlay the
original images from (a) onto the UMAP coordinates (b). The correspondence
between image filenames and the rows of the features in (b) is given by
[this
array](https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week13/exercises/fnames.npy?raw=true).
That is, the $i^{th}$ element of this array is the source image for the
$i^{th}$ row of the features matrix.

```{r, fig.width = 10, fig.height = 10}
s <- .1
for (i in seq_len(nrow(coords))) {
  print(i)
  im <- brick(file.path(data_dir, coords$file[i]))
  p <- p + annotation_raster(as.array(im) / 255, xmin = coords$UMAP1[i] - s, xmax = coords$UMAP1[i] + s, ymin = coords$UMAP2[i] - s, ymax = coords$UMAP2[i] + s)
}

p

#ggsave("~/Downloads/image.png", width = 12, height = 12, dpi = 500)
```



