---
author: ""
title: "Week 13M"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

[Representation Analysis for CIFAR10] This problem asks you to investigate the
features learned by a deep learning model trained to the [CIFAR10
dataset](https://www.cs.toronto.edu/~kriz/cifar.html). We will study the
structure of the dataset and develop some intuition for what representations
different neurons are learning.

a. The code below loads the CIFAR training dataset. Training images and labels
can be accessed using `cifar$x` and `cifar$y`, respectively. Plot the first 15
examples from this dataset and print out their class labels. What does class 7
seem to correspond to?

```{r, echo = TRUE, include = TRUE}
library(keras)
cifar <- dataset_cifar10()$train
```
```{r}
for (i in 1:15) {
  plot(as.raster(cifar$x[i,,,], max = 255))
}

library(tidyverse)
cifar$x[1:15,,,] %>%
  array_tree(1) %>%
  map(~ plot(as.raster(., max = 255)))

cifar$y[1:15]
```






b. We have trained a small model to this dataset, available at this
[link](https://uwmadison.box.com/shared/static/c9kkxyrjb9myfj5knj5vnx0j8gn1uh0s.h5).
Extract and visualize the feature activations associated with the first five
features in layer 1 (`conv2d_8`) for the first image in the dataset.

```{r, echo = TRUE, include = TRUE}
f <- tempfile()
download.file("https://uwmadison.box.com/shared/static/c9kkxyrjb9myfj5knj5vnx0j8gn1uh0s.h5", f)
model <- load_model_hdf5(f)
```
```{r}
activation_model <- keras_model(inputs = model$input, outputs = model$layers[[1]]$output)
features <- activation_model(cifar$x[1:15,,, ])
dim(features)
```

c. The block below extracts feature activations associated with layer 7, for the
first 10% of the dataset. Specifically, the $ij^{th}$ element of `features`
gives the activation of neuron $j$ (in layer 7) on image $i$. Visualize the 10
images that have the highest activation for neuron 1. What does this neuron seem
to be responsive to?
    
```{r, echo = TRUE, include = TRUE}
activation_model <- keras_model(inputs = model$input, outputs = model$layers[[7]]$output)
features <- activation_model(cifar$x[1:5000,,, ]) %>%
  as.matrix()
```


```{r}
which.max(features[,1])
plot(as.raster(cifar$x[456,,,], max = 255))
```


```{r}
library(superheat)
superheat(log(1 + features), pretty.order.rows = TRUE, pretty.order.cols = TRUE)
```

