---
title: "STAT 479 Exercises"
output: rmarkdown::pdf_document
header-includes:
    - \usepackage {hyperref}
    - \hypersetup {colorlinks = true, linkcolor = red, urlcolor = red}
---

```{r, echo = FALSE}
library(knitr)
opts_chunk$set(cache = TRUE, message = FALSE, warning = FALSE, echo = TRUE, include = TRUE)
```


## Module 1: Building Blocks of Data Visualization

1. [Ikea Furniture] The dataset below shows prices of pieces of Ikea furniture.
We will compare prices of different furniture categories and label the
(relatively few) articles which cannot be bought online.

    ```{r}
    library(readr)
    ikea <- read_csv("https://uwmadison.box.com/shared/static/iat31h1wjg7abhd2889cput7k264bdzd.csv")
    ```
    a. Make a plot that shows the relationship between the `category` of furniture
and `price` (on a log-scale). Show each `item_id` as a point -- do not aggregate
to boxplots or ridgelines -- but make sure to jitter and adjust the size the
points to reduce the amount of overlap. _Hint: use the `geom_jitter` layer._
    b. Modify the plot in (a) so that categories are sorted from those with highest
to lowest average prices.
    c. Color points according to whether they can be purchased online. If they
cannot be purchased online, add a text label giving the name of that item of
furniture.

2. [Penguins] The data below measures properties of various Antarctic penguins.
    ```{r, echo = TRUE}
    penguins <- read_csv("https://uwmadison.box.com/shared/static/ijh7iipc9ect1jf0z8qa2n3j7dgem1gh.csv")
    ```
    Create a single plot that makes it easy to answer both of these questions,
        * How is bill length related to bill depth within and across species? 
        * On which islands are which species found?
    
    Read about [Simpson's paradox](https://en.wikipedia.org/wiki/Simpson%27s_paradox) and summarize it in your own words. Then, explain how part (i) provides a real-world example of this paradox.

3. [London Olympics] The data at this
[link](https://uwmadison.box.com/s/rzw8h2x6dp5693gdbpgxaf2koqijo12l) describes
all participants in the London 2012 Olympics.
    a. Create a layered display that shows (i) the ages of athletes across sports
and (ii) the average age within each sport. Use different marks for participants
and for averages. To avoid overplotting, jitter the `Age` variable.
    b. Sort the sports from lowest to highest average age.

4. [German Traffic] We will use `geom_ridgeline` to compare traffic volumes
across a set of cities. The data are read in below,
    ```{r}
    traffic <- read_csv("https://uwmadison.box.com/shared/static/x0mp3rhhic78vufsxtgrwencchmghbdf.csv")
```
    Each row is a timepoint of traffic within a city in Germany. 

    a. Using `geom_ridges`, make a plot of traffic over time, within each of the cities.
    b. Describe two takeaways from the resulting figure. Discuss, but do not implement
    c. Propose an alternative encoding of the same data. What are the strengths of your approach? What are the strengths of the ridgeline approach?

5. [Language Learning] This problem will look at a simplified version of the 
data from the study _A critical period for second language acquisition: Evidence from 2/3 million English speakers_, 
which measured the effect of the the age of initial language learning on
performance in grammar quizzes. We have downloaded the raw data from the
supplementary material and reduced it down to the average and standard
deviations of test scores within (initial learning age) $\times$ (current
age-group) combinations. We have kept a column `n` showing how many participants
were used to compute the associated statistics. The resulting data are available
[here](https://uwmadison.box.com/shared/static/m53dea9w5ipczs3d7nqdnxqxx5ao501b.csv).

    a. Using a tidy approach, define two new fields, `low` and `high`, giving
    confidence intervals for the means in each row. That is, derive new variables 
    according to $\hat{x} \pm 2 * \frac{1}{\sqrt{n}}\hat{\sigma}$.
    
    b. Create a ribbon plot showing confidence intervals for average test scores as
    a function of starting age. Include a line for the average score within that
    combination. Interpret the results of the study.
    
6. [Homelessness] Take a static screenshot from any of the visualizations in
this
[article](https://www.theguardian.com/us-news/ng-interactive/2017/dec/20/bussed-out-america-moves-homeless-people-country-study),
and deconstruct its associated visual encodings. 

    a. What do you think was the underlying data behind the current view? What where
    the rows, and what were the columns?
    b. What were the data types of each of the columns?
    c. What encodings were used? How are properties of marks on the page derived
    from the underlying abstract data?
    d. Is multi-view composition being used? If so, how?
  
7. [Plant Growth Experiment] This problem will give you practice with tidying a
dataset so that it can be easily visualized. The data describe the height of
several plants measured every 7 days. The plants have been treated with
different amounts of a growth stimulant. The first few rows are printed below --
`height.x` denotes the height of the plant on day `x`.
    
    ```{r}
    plants <- read.csv("https://uwmadison.box.com/shared/static/qg9gwk2ldjdtcmmmiropcunf34ddonya.csv")
    ```
    
    a. Propose an alternative arrangement of rows and columns that conforms to the
    tidy data principle.
    
    b. Implement your proposed arrangement from part (a).
    
    c. Using the dataset from (b), design and implement a visualization showing the
    growth of the plants over time according to different treatments.
    
8. [California Wildfires] In this problem, we will interactively visualize a
[dataset](https://uwmadison.box.com/shared/static/k5vvekf1bhh9e16qb9s66owygc70t7dm.csv)
giving statistics of recent California wildfires. The steps below guide you
through the process of building this visualization.
    
    a. (Static) Plot the day of the year that each fire started against the county
    within which it was located. Use the size of the mark to encode the number of
    acres burned in each fire. Sort the counties according to the average latitude
    of the fires within it.
    
    b. (Interactive) Add a tooltip so that the name of the fire can be identified by
    hovering over the points. Introduce a slider to interactively highlight selected
    years.
    
    c. Discuss two takeaways from the visualization. Is there additional information
    that is not described by this visualization or dataset that you think would
    enrich your interpretation?
  
9. [Pokemon]  This problem gives practice in deriving new variables to improve a
faceted plot. The data below give attack and defense statistics for Pokemon,
along with their types. We will build a visualization to answer the question --
how do the different types of Pokemon vary in their attack and defense
potential?

    ```{r}
    pokemon <- read_csv("https://uwmadison.box.com/shared/static/hf5cmx3ew3ch0v6t0c2x56838er1lt2c.csv")
    ```
        
    a. Derive a new column containing the attack-to-defense ratio, defined as
    $\frac{\text{Attack}}{\text{Defense}}$.
    
    b. For each `type_1` group of Pokemon, compute the median attack-to-defense
    ratio.
    
    c. Plot the attack vs. defense scores for each Pokemon, faceted by `type_1`. Use
    the result of (b) to ensure that the panels are sorted from types with highest
    to lowest attack-to-defense ratio.
    
    d. Propose, but do not implement, a visualization of this dataset that makes use
    of dynamic queries. What questions would the visualization answer? What would be
    the structure of interaction, and how would the display update when the user
    provides a cue?
  
9.  In this problem, we'll create a visualization to dynamically query a
[dataset](https://uwmadison.box.com/shared/static/zi72ugnpku714rbqo2og9tv2yib5xped.csv)
of Airbnb rentals in Manhattan in 2019. The steps below guide you through the
process of building this visualization.
    
    a. Make a scatterplot of locations (Longitude vs. Latitude) for all the rentals,
    colored in by `room_type`.
    
    b. Make a histogram of the log-prices^[We've applied a log-transform to the
    original rental prices because otherwise the prices are highly skewed.], with
    stacked colors to distinguish between room types. Vertically concatenate this
    histogram with the scatterplot from (a).
    
    c. Implement a dynamic query so that brushing over price ranges highlights the
    associated rentals on the map.
    
    d. Comment on the resulting visualization. If you had a friend who was
    interested in renting an Airbnb in NYC, what would you tell them?
    
## Module 2: Visualizing Structured Data

1. [Matching Autocorrelation Functions] The purpose of this problem is to build
further intuition about auto-correlation. We'll simulate data with known
structure and then see what happens to the associated autocorrelation functions.
    
    a. The code below simulates a sinusoidal pattern over the course of 2020. Extend
    the code so that `date` and `y` are contained in a tsibble object.
    
    ```{r}
    library(lubridate)
    date <- seq(from = as_date("2020-01-01"), to = as_date("2020-12-31"), by = 1)
    x <- seq(0, 12 * 2 * pi, length.out = length(date))
    y <- sin(x) + rnorm(length(x), 0, .4)
    ```
    
    b. Using the tsibble object, calculate and visualize the induced autocorrelation
    function. Use a maximum lag of 50, and interpret the resulting plot.
    
    c. Write a function to simulate a version of the tsibble above, but with a
    linear trend from 0 to `z`, where `z` is an argument to the function.
    
    d. Using the function from (c), generate 5 datasets with linear trends of
    varying magnitudes. Plot the associated autocorrelation functions and comment on
    the relationship between the strength of the trend and the shape of the
    function.
  
2. [Spotify Time Series]  In this problem, we will study music streaming on
Spotify in 2017. We'll start by looking at some characteristics of the most
streamed song, and then will practice how to extract features from across the
collection of most streamed songs.
    
    a. Let's look at the most streamed song of 2017, which was "Shape of You." The
    dataset
    [here](https://uwmadison.box.com/shared/static/hvplyr3jy6vbt7s80lqgfx81ai4hdl0q.csv)
    contains the number of streams for this song across regions, for each day in
    which it was in the Spotify 100 most streamed songs for that region. Create a
    `tsibble` object from this dataset, keying by `region` and indexing by `date`.
    
    b. Filter to `region == "global"`, and make a `gg_season` plot by month. Comment
    on the what you see.
    
    c. Provide a scatterplot showing the relationship between the number of streams
    of this song in the US and in Canada. Do the same between the US and Japan.
    Briefly comment. **Hint**: Use `pivot_wider` to spread the time series for each
    region across columns of a
    [reshaped](https://krisrs1128.github.io/stat479/posts/2021-01-27-week4-2/)
    dataset.
    
    d. The dataset
    [here](https://uwmadison.box.com/shared/static/xj4vupjbicw6c8tbhuynw0pll6yh1w0d.csv)
    contains similar data, but for all songs that appeared in the Spotify 100 for at
    least 200 days in 2017. We have filtered to only the global totals. Read these
    data into a tibble, keyed by `artist:region` and extract features of the
    `streams` time series using the `features` function in the feasts library. It is
    normal to see a few errors reported by this function, it just means that some of
    the statistics could not be calculated.
    
    e. Which tracks had the highest and lowest `trend_strength`'s? Visualize their
    streams over the course of the year.
    
3. [NYC Trees] In this problem, we'll use vector data to enrich a visualization
of trees in New York City. In the process, we'll practice reading in and
generating summaries of geospatial data.
    
    a. The data at this
    [link](https://uwmadison.box.com/shared/static/t1mk6i4u5ks5bjxaw2c7soe2z8i75m2o.csv)
    include a subset of data from the New York City Tree Census. Make a scatterplot
    of the locations of all trees in the data, coloring in by tree species group and
    faceting by health.
    
    b. Suppose we wanted to relate these data to characteristics of the built
    environment. We have curated public data on
    [roads](https://uwmadison.box.com/shared/static/28y5003s1d0w9nqjnk9xme2n86xazuuj.geojson)
    and
    [buildings](https://uwmadison.box.com/shared/static/qfmrp9srsoq0a7oj0e7xmgu5spojr33e.geojson)
    within the same neighborhood. Read these data into `sf` objects using `read_sf`.
    For both datasets, report (i) the associated CRS and (ii) the geometry type
    (i.e., one of point, linestring, polygon, multipoint, multilinestring,
    multipolygon, geometry collection).
    
    c. Generate a version of the plot in (a) that has the roads and buildings in the
    background.
    
4. [Himalayan Glaciers] In this problem, we'll apply the reading's discussion of
raster data to understand a
[dataset](https://uwmadison.box.com/shared/static/2z3apyg4t7ct5qd4mcwh9rpr63t02jql.tif)
containing Landsat 7 satellite imagery of a Himalayan glacier system.
    
    a. Read the data into a `brick` object. What is the spatial extent of the file
    (that is, within what geographic coordinates do we have data)? How many layers
    of sensor measurements are available?
    
    b. Generate an RGB image of this area. In Landsat 7, the first three layers (B1,
    B2, and B3) provide the red, green, and blue channels.
    
    c. Make a plot of the slopes associated with each pixel within this region.
  
5. [CalFresh Enrollment] In this problem, we will investigate spatial and
temporal aspects of enrollment in CalFresh, a nutritional assistance program in
California.
    
    a. The code below reads in the CalFresh data. We've filtered out February 2019,
    since benefits were distributed differently in this month, leading to outliers
    for most counties. Extract features of the `calfresh` time series using the
    `features` function in the feasts library.
    
    ```{r}
    library(tsibble)
    calfresh <- read_csv("https://uwmadison.box.com/shared/static/rduej9hsc4w3mdethxnx9ccv752f22yr.csv") %>%
      filter(date != "2019 Feb") %>%
      mutate(date = yearmonth(date)) %>%
      as_tsibble(key = county, index = date)
    ```
    
    b. Visualize CalFresh enrollment over time for the counties with the highest and
    lowest `seasonal_strength_year`.
    
    c. The code below reads in a vector dataset demarcating the county boundaries in
    California. Join in the features dataset from (a) with this these vector data.
    Use this to produce a map with each county shaded in by its
    `seasonal_strength_year`.
    
    ```{r}
    library(sf)
    counties <- read_sf("https://uwmadison.box.com/shared/static/gropucqxgqm82yhq13do1ws9k16dnxq7.geojson")
    ```
    
    d. Propose, but do not implement, a visualization of this dataset that makes use
    of dynamic queries. What questions would the visualization answer? What would be
    the structure of interaction, and how would the display update when the user
    provides a cue?
    
6. [Political Book Recommendations] In this problem, we'll study a network
dataset of Amazon bestselling US Politics books. Books are linked by an edge if
they appeared together in the recommendations ("customers who bought this book
also bought these other books").
    
    a. The code below reads in the edges and nodes associated with the network. The
    edges dataset only contains IDs of co-recommended books, while the nodes data
    includes attributes associated with each book. Use the edges dataset to create
    an igraph graph object, and use the `ggnetwork` function to construct a
    data.frame summarizing the layout of the network. Use a `layout_with_fr` layout,
    as in the reading.
    
    ```{r}
    edges <- read_csv("https://uwmadison.box.com/shared/static/54i59bfc5jhymnn3hsw8fyolujesalut.csv")
    nodes <- read_csv("https://uwmadison.box.com/shared/static/u2x392i79jycubo5rhzryxjsvd1jjrdy.csv", col_types = "ccc")
    ```
    
    b. The output from (a) does not include any attributes about the books, since
    this is only available in the `nodes` dataset, and we built the graph layout
    using only `edges`. Join in the node attribute data. **Hint**: Use `left_join`, but
    using the `by` argument to ensure that `name` in the output of `ggnetwork` and
    `Id` in the original `nodes` dataset are associated with one another.
    
    c. Use the result from part (b) to visualize the network using. Include the
    book's title in the node label, and shade in the node according to political
    ideology.
  
## Module 3: Visualizing Models

1. [Polio incidence] In this problem, we will use a heatmap to visualize a large
collection of time series. The
[data](https://uwmadison.box.com/s/nm7yku4y9q7ylvz5kbxya3ouj2njd0x6), prepared
by the [Tycho Project](https://www.tycho.pitt.edu/), contain weekly counts of
new polio cases in the United States, starting as early as 1912 (for some
states).
    
    a. Pivot the raw dataset so that states appear along rows and weeks appear along
    columns. Fill weeks that don't have any cases with 0's.
    
    ```{r}
    polio <- read_csv("https://uwmadison.box.com/shared/static/nm7yku4y9q7ylvz5kbxya3ouj2njd0x6.csv")
    ```
    
    b. Use the `superheat` package to make a heatmap of the data from (a). Have the
    color of each tile represent `log(1 + cases)`, rather than the raw counts.
    Reorder the states using a hierarchical clustering by setting `pretty.order.rows
    = TRUE`.
    
    c. Supplement the view from part (b) with a barchart showing the US total
    incidence during every given week. Interpret the resulting visualization.
    *Hint: use the `yt` argument of superheat.*
    
    d. Describe types of annotation would improve the informativeness of the plot
    made in part (c). Also, describe one advantage and one disadvantage of
    visualizing a collection of time series as a heatmap, instead of as a collection
    of line plots.
    
2. [Silhouette Statistics Simulation] This problem uses simulation to build
intuition about silhouette statistics. The function below simulates a mixture of
three Gaussians, with means evenly spaced out between `(start, 0)` and `(end,
0)`. We will investigate what happens to the silhouette statistics for each
point as the three clusters are made to gradually overlap.
    
    ```{r, fig.width = 6, fig.height = 2, fig.cap = "An example simulated dataset from problem 2." }
    library(tibble)
    library(purrr)
    library(ggplot2)
    
    mixture_data <- function(start = 0, end = 10, K = 3, n = 100) {
      mu <- cbind(seq(start, end, length.out = K), 0)
      map_dfr(1:K, ~ MASS::mvrnorm(n, mu[., ], diag(2)) %>% as_tibble())
    }
    
    ggplot(mixture_data()) +
      geom_point(aes(V1, V2)) +
      coord_fixed() +
      labs(x = "x", y = "y")
    ```
    
    a. Simulate versions of these data where the `end` argument is decreased from 10
    to 0.
    
    b. Write a function that performs $K$-means and computes the silhouette
    statistic given any one of the datasets generated in part (a). Use this function
    to compute the silhouette statistics for each point in the simulated datasets.
    
    c. Visualize the silhouette statistics from part (b) overlaid onto the simulated
    data. Discuss the results.
    
3. [Taxi Trips] In this problem, we will use hierarchical clustering to find
typical taxi trip trajectories in Porto, Portugal. The data are a subset from
the the [ECML / PKDD 2015
Challenge](http://www.geolink.pt/ecmlpkdd2015-challenge/dataset.html) -- the
link provides a complete data dictionary. We have preprocessed it into two
formats. The first
([`wide`](https://uwmadison.box.com/shared/static/cv0lij4d9gn3s8m2k98t2ue34oz6sbj5.csv))
includes each taxi trip on its own row, with latitude and longitude coordinates
along the journey given as separate columns (`x_0, y_0` is the origin of the
trip and `x_15`, `y_15` is the destination). The second
([`long`](https://uwmadison.box.com/shared/static/098cjaetm8vy0mufq21mc8i9nue2rr2b.csv))
format spreads each point of the journey into a separate row.
    
    a. Filter the `long` form of the data down to trip `1389986517620000047` and
    plot its trajectory as a sequence of points.
    
    ```{r}
    trips <- read_csv("https://uwmadison.box.com/shared/static/098cjaetm8vy0mufq21mc8i9nue2rr2b.csv", col_types = cols(TRIP_ID = col_character()))
    trips_wide <- read_csv("https://uwmadison.box.com/shared/static/cv0lij4d9gn3s8m2k98t2ue34oz6sbj5.csv", col_types = cols(TRIP_ID = col_character()))
    ```
    
    b. We could hierarchically cluster rows in either the `wide` or the `long`
    format datasets. How would the interpretation of the results differ between the
    two approaches?
    
    c. Compute a hierarchical clustering of the `wide` format data, using only the
    columns starting with `x` or `y` as features.
    
    d. Cut the hierarchical clustering tree so that 8 clusters are produced.
    Visualize the trajectories of the taxi trips either colored or faceted by their
    cluster.
    
4. [Food nutrients] This problem will use PCA to provide a low-dimensional view
of a 14-dimensional nutritional facts
[dataset](https://uwmadison.box.com/shared/static/nmgouzobq5367aex45pnbzgkhm7sur63.csv).
The data were originally curated by the USDA and are regularly used in
[visualization studies](http://bl.ocks.org/syntagmatic/raw/3150059/).
    
    ```{r}
    nutrients <- read_csv("https://uwmadison.box.com/shared/static/nmgouzobq5367aex45pnbzgkhm7sur63.csv")
    ```
    
    a. Define a tidymodels `recipe` that normalizes all nutrient features and
    specifies that PCA should be performed.
    
    b. Visualize the top 6 principal components. What types of food do you expect to
    have low or high values for PC1 or PC2?
    
    c. Compute the average value of PC2 within each category of the `group` column.
    Give the names of the groups sorted by this average.
    
    d. Visualize the scores of each food item with respect to the first two
    principal components. Facet the visualization according to the `group` column,
    and sort the facets according to the results of part (c). How does the result
    compare with your guess from part (b)?
    
6. [Topics in _Pride and Prejudice_] This problem uses LDA to analyze the full
text of _Pride and Prejudice_. The code below creates two R objects, `paragraph`
and `dtm`. `paragraph` is a data.frame whose rows are paragraphs^[We've filtered
very short paragraphs; e.g., from dialogue.] from the book. `dtm` is a
`DocumentTermMatrix` containing word counts across the same paragraphs -- the
$i^{th}$ row of `dtm` corresponds to the $i^{th}$ row of `paragraph`.
    
    ```{r}
    library(readr)
    library(dplyr)
    library(tidytext)
    
    paragraphs <- read_csv("https://uwmadison.box.com/shared/static/pz1lz301ufhbedzsj9iioee77r95xz4v.csv") 
    dtm <- paragraphs %>%
      unnest_tokens(word, text) %>%
      filter(!word %in% stop_words$word) %>%
      count(paragraph, word) %>%
      cast_dtm(paragraph, word, n)
    ```
    
    a. Fit an LDA model to `dtm` using 6 topics. Set the seed by using the argument
    `control = list(seed = 479)` to remove any randomness in the result.
    
    b. Visualize the top 30 words within each of the fitted topics. Specifically,
    create a faceted bar chart where the heights of the bars correspond to word
    probabilities and the facets correspond to topics. Reorder the bars so that each
    topic's top words are displayed in order of decreasing probability.
    
    c. Find the paragraph that is the purest representative of Topic 2. That is, if
    $\gamma_{ik}$ denotes the weight of topic $k$ in paragraph $i$, then print out
    paragraph $i^{\ast}$ where $i^{\ast} = \arg \max_{i}\gamma_{i2}$. Verify that
    the at least a few of the words with high probability for this topic appear.
    Only copy the first sentence into your solution.
  
7. [Single-Cell Genomics]  In this problem, we will apply UMAP to a
[dataset](https://uwmadison.box.com/shared/static/ai539s30rjsw5ke4vxbjrxjaiihq7edk.csv)
of Peripheral Blood Mononuclear Cells (PBMC) released by 10X Genomics. The first
column, `cell_tag`, gives an identifier for each cell in the dataset. All other
columns are molecules that were detected in that cell. For example, CD74 is a
molecule often found on the surface of T-cells.
    
    a. Define a tidymodels `recipe` that specifies that UMAP should be performed
    with the parameters `learn_rate = 0.1` and `neighbors = 5`. There is no need to
    normalize these data, as they have been normalized in advance using methods
    tailored to single-cell genomics data.
    
    ```{r}
    pbmc <- read_csv("https://uwmadison.box.com/shared/static/ai539s30rjsw5ke4vxbjrxjaiihq7edk.csv")
    ```
    
    b. Compute the UMAP embeddings across cells. Color points in by their value of
    the GNLY molecule.
  
8. [Political Blogs] In this problem, we will use topic modeling to see how the topics discussed
across the political blogsphere evolved over course of 2008. We have fitted a
model to a subsample of [The CMU 2008 Political Blog
Corpus](http://reports-archive.adm.cs.cmu.edu/anon/ml2010/CMU-ML-10-101.pdf),
and the resulting topicmodels object is
[here](https://uwmadison.box.com/shared/static/nen0mdsahjpa4d67g9choq5y481ws4mk.rda).
    
    a. Use the discrepancy formula described in lecture,
    \begin{align*}
    D\left(k, l\right) := \beta_{kw}\log\left(\frac{\beta_{kw}}{\beta_{lw}}\right) + \left(\beta_{lw} - \beta_{kw}\right)
    \end{align*}
    to identify 50 words that can be used to discriminate between the fitted topics.
      
    b. Make a heatmap that displays the $\beta_{kw}$ probabilities of the words in part (b).
      
    c. Using the document [metadata
    file](https://uwmadison.box.com/shared/static/qg91329100k9ah11jvvkvbcy5dy7adnw.csv),
    design and implement a visualization that shows variation in the memberships
    $\gamma_{dk}$ either over time, across blogs, or in relation to political
    leaning. Suggest some interpretations for patterns that you see.

9. [A Bikesharing Model] In this and the next problem, we will visualize models
fitted to predict bikesharing demand in a subset of the Capitol Bikesharing
[dataset](https://uwmadison.box.com/shared/static/aa91qdqehagag8wg8mqsm4z5b4g2hu0x.csv).
We will see what types of features are learned by different types of models,
whether there are any interactions between features, and whether linear and
nonlinear approaches are substantively different.

    a. The code below fits a gradient boosting machine (GBM) to predict bikesharing
    demand (count) using all features available, except day of the year. Visualize
    the Ceteris Paribus profiles for the temperature and humidity variables, and
    provide a brief interpretation.
          
    ```{r, echo = TRUE, include = TRUE}
    bike <- read_csv("https://uwmadison.box.com/shared/static/aa91qdqehagag8wg8mqsm4z5b4g2hu0x.csv")
    x <- dplyr::select(bike, -count, -dteday)
    hyper <- data.frame(n.trees = 100, interaction.depth = 4, shrinkage = 0.1, n.minobsinnode = 10)
    fit <- train(x = x, y = bike$count, method = "gbm", tuneGrid = hyper, verbose = FALSE)
    ```
    
    b. Has the GBM learned an interaction effect between the hour of day (`hr`) and
    weekend (`weekend`) features? Briefly justify your answer using a grouped CP
    plot.
  
10. [Contrastive Profiles for Bikesharing] This problem continues the exploration
in the previous one. Here, we study whether the choice of GBM was critical, or
whether any other model would have learned the same relationship between the
predictors and bikesharing demand.

    a. The code below fits a lasso model to the same input and output dataset.
    Provide contrastive partial dependence profiles between the lasso and the GBM
    from above, focusing on the hour (`hr`), humidity (`hum`), and temperature
    (`temp`) features. Comment on the result.

    ```{r, echo = TRUE, include = TRUE}
    hyper <- data.frame(lambda = 0.001, alpha = 0)
    fit_lm <- train(x = x, y = bike$count, method = "glmnet", tuneGrid = hyper)
    ```
        
    b. Repeat part (a), but comparing the GBM with the CART model fitted below.
    Comment on the result.
  
```{r, echo = TRUE, include = TRUE}
library(caret)
hyper <- data.frame(cp = 0.01)
fit_cart <- train(x = x, y = bike$count, method = "rpart", tuneGrid = hyper)
```

11. [Representation Analysis for CIFAR10] This problem asks you to investigate
the features learned by a deep learning model trained to the [CIFAR10
dataset](https://www.cs.toronto.edu/~kriz/cifar.html). We will study the
structure of the dataset and develop some intuition for what representations
different neurons are learning.

    a. The code below loads the CIFAR training dataset. Training images and labels
    can be accessed using `cifar$x` and `cifar$y`, respectively. Plot the first 15
    examples from this dataset and print out their class labels. What does class 7
    seem to correspond to?
  
    ```{r, echo = TRUE, include = TRUE}
    library(keras)
    cifar <- dataset_cifar10()$train
    ```
    
    b. We have trained a small model to this dataset, available at this
    [link](https://uwmadison.box.com/shared/static/c9kkxyrjb9myfj5knj5vnx0j8gn1uh0s.h5).
    Extract and visualize the feature activations associated with the first five
    features in layer 1 (`conv2d_8`) for the first image in the dataset.
    
    ```{r, echo = TRUE, include = TRUE}
    f <- tempfile()
    download.file("https://uwmadison.box.com/shared/static/c9kkxyrjb9myfj5knj5vnx0j8gn1uh0s.h5", f)
    model <- load_model_hdf5(f)
    ```
  
    c. The block below extracts feature activations associated with layer 7, for the
    first 10% of the dataset. Specifically, the $ij^{th}$ element of `features`
    gives the activation of neuron $j$ (in layer 7) on image $i$. Visualize the 10
    images that have the highest activation for neuron 1. What does this neuron seem
    to be responsive to?
        
  ```{r, echo = TRUE, include = TRUE}
  activation_model <- keras_model(inputs = model$input, outputs = model$layers[[7]]$output)
  features <- activation_model(cifar$x[1:5000,,, ]) %>%
    as.matrix()
  ```

12. [Household Radon Levels] This problem analyzes a
[dataset](https://uwmadison.box.com/shared/static/3yn994tc1ft73z3br4ys18uri700gvik.csv)
of household radon levels in Minnesota. The data were gathered by the EPA in [an
effort](https://pubmed.ncbi.nlm.nih.gov/8919076/) to understand which factors
lead to increased exposure to this carcinogenic gas; e.g., it is known that
radon levels are higher in basements than on the ground floor. A secondary goal
of this problem is to illustrate the use of the `rstanarm` and `tidybayes`
packages to extract prior and posterior predictive distributions. While these
packages cannot be applied as generally as the methods discussed in lecture,
they provide an elegant interface for specific types of Bayesian models. You
will not need to write any code using these packages -- we only ask that you
visualize their outputs.
    
    a. Design and implement a visualization that describes the variation in
    `log_radon` levels from county to county.
  
    ```{r, echo = TRUE, include = TRUE}
    radon <- read_csv("https://uwmadison.box.com/shared/static/3yn994tc1ft73z3br4ys18uri700gvik.csv")
    ```
      
    b. The code below simulates prior predictive data from two potential Bayesian
    models of `log_radon`. Both models have the form `log_radon ~ (1 | county) +
    floor`, which means each county - floor level combination gets its own mean. In
    the first model, a vague $\mathcal{N}\left(0, 1000\right)$ prior is used, while
    in the second, a somewhat more informative $\mathcal{N}\left(0, 10\right)$ is
    used instead.
    
    ```{r, echo = TRUE, include = TRUE}
    library(rstanarm)
    priors <- list(
      "vague" = stan_lmer(log_radon ~ (1 | county) + floor, prior = normal(0, 1000), data = radon, prior_PD = 1, refresh = 0),
      "informative" = stan_lmer(log_radon ~ (1 | county) + floor, prior = normal(0, 10), data = radon, prior_PD = 1, refresh = 0)
    )
    ```
          
    The code below extracts 5 simulated datasets from the prior predictive. Each
    row is a simulated house; the simulation number is denoted by `.draw`. The
    index to the original dataset is given by `.row`. The simulated `log_radon`
    levels are given in the `.prediction` column. The `prior` column describes
    which type of prior was used for that simulation.
     
    ```{r, echo = TRUE, include = TRUE}
    prior_preds <- map_dfr(priors, ~ add_predicted_draws(radon, .), .id = "prior") %>%
      filter(.draw > 3995)
    
    head(prior_preds)
    ```
        
    Design and implement a visualization to compare the two types of priors. Which
    prior seems more plausible?
      
    c. The code below draws posterior predictive samples from two new models, which
    either ignore or model variation from county to county (`lm` and `hierarchical`
    respectively). Both use information about whether the measurement came from a
    basement. Only the last 10 simulated runs are kept. Using these data, make
    boxplots of posterior predictive samples across counties, for each of the two
    models. What are the main differences between the two models? In what ways are
    the county level effects (as you studied in part (a)) accurately or
    inaccurately modeled?
     
    ```{r, echo = TRUE, include = TRUE}
    library(tidybayes)
    posterior_preds <- list(
      "lm" = stan_glm(log_radon ~ floor, prior = normal(0, 10), data = radon, chains = 1, refresh = 0),
      "hierarchical" = stan_lmer(log_radon ~ (1 | county) + floor, prior = normal(0, 10), data = radon, chains = 1, refresh = 0)
    ) %>%
      map_dfr(~ add_predicted_draws(radon, .), .id = "model") %>%
      filter(.draw > 990)
    ```
  
13. [Reading Response] This problem asks you to reflect on one of the three
readings from week 14
([a](https://www.datasketch.es/project/655-frustrations-doing-data-visualization),
[b](https://simplystatistics.org/2019/04/17/tukey-design-thinking-and-better-questions/),
[c](https://datavis.ca/papers/hbook.pdf)). Prepare a brief (1 paragraph)
response. For example, you may discuss any of the following points,
    
    * Are there points from the reading that you strongly agree or disagree with?
    * Are there lessons from the reading that you think you might incorporate into
    your future projects or plans?
    * How would you explain the main ideas of the reading to a friend who is not
    technically trained in visualization?
  