---
author: "Kris Sankaran"
title: "Week6W"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
library(shiny)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = FALSE, dpi = 200, fig.align = "center", out.width = 650, fig.height = 3, fig.width = 9)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Time Series Visualization (Part 2)

.pull-left[
  March 4, 2022
]
 
---

### Announcements

* Portfolio 2 - March 13
* Project Milestone 2 - March 13
* Can pick up exam after class or at office hours
  
---

### Today

By the end of the class, you should be able to...

* Extract features from a tsibble object
* Navigate a collection of time series based on their features

---

### Exercise 6.1a-b review

In this problem, we will study music streaming on Spotify in 2017. We'll start
by looking at some characteristics of the most streamed song, and then will
practice how to extract features from across the collection of most streamed
songs.
    
a. Let's look at the most streamed song of 2017, which was "Shape of You." The
dataset
[here](https://uwmadison.box.com/shared/static/hvplyr3jy6vbt7s80lqgfx81ai4hdl0q.csv)
contains the number of streams for this song across regions, for each day in
which it was in the Spotify 100 most streamed songs for that region. Create a
`tsibble` object from this dataset, keying by `region` and indexing by `date`.

b. Filter to `region == "global"`, and make a `gg_season` plot by month. Comment
on the what you see.

---

`index` gives the time index, and `key` says that we should expect a different
time series for each region.

```{r}
library(tidyverse)
library(tsibble)
songs <- read_csv("https://uwmadison.box.com/shared/static/hvplyr3jy6vbt7s80lqgfx81ai4hdl0q.csv") %>%
  as_tsibble(index = date, key = region)

head(songs)
```

---

```{r, fig.height= 4, fig.width = 6, out.width = 500}
library(feasts)
songs %>%
  filter(region == "global") %>%
  gg_season(period = "month")
```

---

```{r, fig.height= 4, fig.width = 6, out.width = 500}
library(scales)
cols <- viridis_pal()(10)
songs %>%
  filter(region == "global") %>%
  gg_season(period = "month", pal = cols, size = 1, alpha = 0.8) +
  scale_y_continuous(labels = label_number_si()) +
  scale_x_date(expand = c(0, 0), labels = label_date("%d"))
```

---

### Live Coding

Continuing the carbon emissions example.

```{r}
carbon_ts <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week6/data/cleaned.csv") %>%
  pivot_longer(-c("country", "Country Code"), names_to = "year") %>%
  mutate(year = as.integer(year)) %>%
  filter(country != "Switzerland") %>%
  as_tsibble(index = year, key = country)

head(carbon_ts)
```

---

### Carbon Example

This dataset is a collection of time series. To help navigate them, it helps to
extract summary features about them,

  * The amount of seasonal or cyclic structure
  * The presence of trends
  * The degree of short and long-run autocorrelation

---

### Exercise

* Exercise 6.2 on Canvas
* Can discuss, but submit individually
* Only have to do parts d - e! Try other parts if you have time.