---
author: "Kris Sankaran"
title: "Week6W"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
library(tsibble)
library(feasts)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

```{r}
library(tidyverse)
library(tsibble)
library(feasts)

carbon_ts <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week6/data/cleaned.csv") %>%
  pivot_longer(-c("country", "Country Code"), names_to = "year") %>%
  mutate(year = as.integer(year)) %>%
  filter(country != "Switzerland") %>%
  as_tsibble(index = year, key = country)
```

```{r}
summaries <- carbon_ts %>%
   features(value, features = list(mean = mean, sd = sd))
summaries %>%
  arrange(-mean)
```

```{r}
summaries <- carbon_ts %>%
   features(value, feature_set(tag = "trend"))
summaries
```

y[t] = trend[t] + seasonal[t] + remainder[t]


Q1: Can you give an intuitive interpretation of the following features that are
computed by this package?

  * trend_strength
  * curvature

Hint: You can look either in the
[documentation](https://feasts.tidyverts.org/reference/index.html) for the
package or an associated [book chapter](https://otexts.com/fpp3/stlfeatures.html).

Q2: Can you design a graphical query that lets us look up time series based on
their features?

---

If we try a Shiny example, ...

update the server,

1. Write the observe event to change the selected reactiveValue
2. ggplot2 for the histogram
3. ggplot2 for the time series
4. customize the appearance

```{r}
library(shiny)
trends <- carbon_ts %>%
  features(value, features = feature_set(tags = "stl"))
carbon_trends <- left_join(carbon_ts, trends)

ui <- fluidPage(
  plotOutput("histogram", brush = brushOpts("brush", direction = "x")),
  plotOutput("trends")
)

server <- function(input, output) {
  selected <- reactiveVal(rep(TRUE, nrow(carbon_trends)))
  output$histogram <- renderPlot({
    ggplot(carbon_trends) +
      geom_histogram(aes(spikiness))
  })
  output$trends <- renderPlot({
    carbon_trends %>%
      mutate(selected = selected()) %>%
      ggplot() +
      geom_line(aes(year, value, group = country, col = as.numeric(selected), alpha = as.numeric(selected))) +
      scale_alpha(range = c(0.1, 1))
  })
  observeEvent(
    input$brush, {
      selected(brushedPoints(carbon_trends, input$brush, allRows = TRUE)$selected_)
    }
  )
}

shinyApp(ui, server)
```