---
author: "Kris Sankaran"
title: "Week6W"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = FALSE, dpi = 200, fig.align = "center", out.width = 650, fig.height = 3, fig.width = 9)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

```{r}
library(tidyverse)
library(tsibble)
library(lubridate)
```

```{r}
carbon <- read_csv("data/cleaned.csv")
carbon
```

It will be easier to visualize these data if they are made tidy.
```{r}
carbon <- carbon %>%
  pivot_longer(-c("country", "Country Code"), names_to = "year")
```

This doesn't work, because year is a string.
```{r, eval = FALSE}
as_tsibble(carbon, index = "year", key = "country")
```

This still doesn't work -- there are duplicates!
```{r}
carbon %>%
  mutate(year = as.integer(year)) %>%
  as_tsibble(index = year, key = country)
```

For some reason, there are duplicated entries for Switzerland. This is one of
the reasons for using something like `tsibble`.
```{r}
carbon %>%
  mutate(year = as.integer(year)) %>%
  duplicates(key = country, index = year) %>%
  arrange(year)
```

This code finally makes the tsibble.
```{r}
carbon_ts <- carbon %>%
  filter(country != "Switzerland") %>%
  mutate(year = as.integer(year)) %>%
  as_tsibble(index = year, key = country)
```

This is an example of the sort of plot we can automatically make with tsibble.
If we had used `carbon` instead of `carbon_ts` as the input, the plot would not
have worked.
```{r}
library(feasts)
carbon_ts %>%
  filter(country == "United States") %>%
  gg_lag()
```

Here are some basic kinds of time series plots.
```{r}
ggplot(carbon_ts) +
  geom_line(aes(year, value, group = country))
ggplot(carbon_ts) +
  geom_tile(aes(year, reorder(country, value), fill = log(1 + value)))
```

What if we filter down the countries that are shown?

```{r}
keep_countries <- carbon_ts %>%
  group_by(country) %>%
  filter(mean(value) > 1) %>%
  pull(country)

carbon_ts %>%
  filter(country %in% keep_countries) %>%
  ggplot() +
  geom_tile(aes(year, reorder(country, value), fill = log(1 + value)))
```

I was surprised, but it seems that the Bahamas really does have a spike in
carbon emissions during the 1970s
([source](https://www.google.com/search?q=bahamas+carbon+emissions&rlz=1C5CHFA_enUS955US955&oq=bahamas&aqs=chrome.0.69i59l2j69i57j0i67l4j0i20i263i512j0i512j46i199i465i512.1468j0j7&sourceid=chrome&ie=UTF-8)).


What if we want a smoothed version of the plot? This is a more complicated
operation (I don't expect you to remember this), but sometimes it comes in
handy.

```{r}
library(slider)
smoothed_carbon <- carbon_ts %>%
  split(.$country) %>%
  map_dfr(
    ~ mutate(
        ., 
        smoothed = slide_dbl(value, ~ mean(.), .before = 5, .after = 5)
      )
  ) 

ggplot(smoothed_carbon) +
  geom_line(aes(year, smoothed, group = country))
```

What if we want to be able to hover and get the country names?

```{r}
library(plotly)
p <- ggplot(smoothed_carbon) +
  geom_line(aes(year, smoothed, group = country))
ggplotly(p)
```

