---
author: "Kris Sankaran"
title: "Week6W"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = FALSE, dpi = 200, fig.align = "center", out.width = 650, fig.height = 3, fig.width = 9)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Time Series Visualization (Part 1)

.pull-left[
  March 2, 2022
]
 
---

### Announcements

* Portfolio 2 posted (March 13)
* Can pick up exam after class or at office hours
  
---

### Today

By the end of the class, you should be able to...

* Construct and visualize a `tsibble` time series object
* Discuss the presence of trend, seasonal, or cyclic structure in a time series

---

### Midterm 1 Review

* Quartiles:
* For grades / curve, see syllabus 
  - Best of absolute scores and class ranking
* Q4b and Q7 were most challenging (by design)
  - We were generous with partial credit

---

### Q4b
    
Provide code to compute the proportion of each quarter's sales that came from
each region.

```{r}
sales <- data.frame(
  region =  c('A', 'A', 'A', 'A', 'B', 'B', 'B', 'B'),
  quarter = c('Q1', 'Q2', 'Q3', 'Q4', 'Q1', 'Q2', 'Q3', 'Q4'),
  sales = c(6, 5, 3, 2, 4, 8, 2, 6)
)
sales
```

---

### Q4b

This is tricky because the data we need to get group totals before we can
compute the proportions.

```{r}
sales %>%
  group_by(quarter) %>%
  mutate(total = sum(sales))
```

---

### Q4b

```{r}
sales %>%
  group_by(quarter) %>%
  mutate(
    total = sum(sales),
    prop = sales / total
  ) %>%
  select(region, quarter, prop)
```

---

### Q7

The code below sets up a Shiny app for interactively visualizing athlete weight
and heights in the 2012 London Olympics. We would like to have an interactive
scatterplot of `Weight` vs. `Height, cm` that updates which points (athletes)
are highlighted depending on sports have been selected by a dropdown menu. Code
for generating the scatterplot is provided in the function `scatterplot`.

```{r}
scatterplot <- function(df) {
  ggplot(df) +
    geom_point(
      aes(Weight, `Height, cm`, 
          alpha = as.numeric(selected),
          size = as.numeric(selected))
      ) +
    scale_alpha(range = c(0.05, .8)) +
    scale_size(range = c(0.1, 1))
}
```

---

Provide server code which would allow the scatterplot to update the highlighted
athletes depending on the currently selected sports.

```{r, echo = TRUE}
ui <- fluidPage(
  selectInput("dropdown", "Select a Sport", choices = unique(olympics$Sport), multiple = TRUE),
  plotOutput("scatterplot")
)

server <- function(input, output) {
  ### fill this in...
}
```

---

* There is one output (`scatterplot`), and it needs a `renderPlot`
* We can update the `selected` column depending on the input

```{r, eval = FALSE}
output$scatterplot <- renderPlot({
  olympics %>%
    mutate(selected = Sport %in% input$dropdown) %>%
    scatterplot()
})
```

---

We have been asked to also print a table of the selected athletes. Assume the UI
has the form,
    
```{r}
ui <- fluidPage(
  selectInput("dropdown", "Select a Sport", choices = unique(olympics$Sport), multiple = TRUE),
  plotOutput("scatterplot"),
  dataTableOutput("table") 
)
```
  
Describe changes to your solution to (a) to meet the new requirements. How
would you minimize code duplication? Be as specific as possible.

---

* We need a new output for the table
* We can use a reactive to minimize code duplication

```{r, eval = FALSE}
current_athletes <- reactive({
  olympics %>%
    mutate(selected = Sport %in% input$dropdown)
})

output$scatterplot <- renderPlot(scatterplot(current_athletes()))
output$table <- renderDataTable({
  current_athletes() %>% filter(selected)
})
```

---

## Live Coding Example

---

### Ground Rules

* Take out your laptops, follow-along, and ask questions 
* I will ask for your input for the visualization strategy
* You and I will both run into bugs -- this is a normal part of the process
* For today, see the `week6W-code-starter.Rmd` file

---

### Carbon Dataset

This is a dataset explored by one of the project
[groups](https://observablehq.com/@wang2274/milestone-4-team-4) last year. It is
a dataset carbon emissions at country-by-year granularity, compiled by the World
Bank.

We will,

  * Make a `tsibble` object from these data
  * Try answering some time-series related questions visually

What are some of the sorts of questions we might want to answer?

---

Here are the data.

```{r}
carbon <- read_csv("data/cleaned.csv")
carbon
```

---

### Exercise

* Exercise 6.1 on Canvas
* Can discuss, but submit individually
* Only have to do parts a - b! Try other parts if you have time.
