---
author: ""
title: "Week9M Code Starter"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

1. [Women, Business and the Law] The World Bank's Women, Business, and the Law
program has curated a dataset about gender equality across countries since 1971.
Since there are more than 30 variables summarizing each country for each year,
we will use clustering to observe the general trends across all measures.

    a. The code below reads in the data, simplifies column names (saving the
    original names in the `question_mapping` data.frame), and converts text Yes
    / No responses to numerical 0 / 1. Widen the dataset so that each row
    corresponds to a single country $\times$ year combination.
    
    ```{r}
    # read in and create a codebook
    survey <- read_csv("https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week9/exercises/Viz5-July-2020-economic-empowerement.csv?raw=true")
    question_mapping <- survey %>%
      select(Question, `Question Category`) %>%
      distinct() %>%
      mutate(q_id = str_c("Q", row_number()))
    
    recode_vector <- setNames(question_mapping$Question, question_mapping$q_id)
    survey <- survey %>%
      mutate(
        Question = fct_recode(Question, !!!recode_vector),
        answer = as.numeric(ifelse(`Text Answer` == "Yes", 1, 0))
      ) %>%
      select(Country, `WBL Report Year`, Question, answer)
    
    survey <- survey %>%
      pivot_wider(names_from = "Question", values_from = "answer")
    ```
    
    b. Apply $K$-means to the question responses (you may choose $K$). Visualize
    the centroids and briefly interpret each cluster in context, using the
    original text of the questions stored in `question_mapping`.
    
    c. Visualize change in cluster sizes over time. Provide a brief
    interpretation (for example, in which types of questions is there the most /
    least progress towards equality?).
