---
author: ""
title: "Week9W Code Starter"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

1. [Beijing Air Pollution] This problem studies daily air pollution and weather
patterns in Beijing from 2010 to 2014. Each row in this dataset gives
measurements for three variables, `pollution`, `temp`, and `wndspd` over time
for a single day, indexed by `date`. For example the `temp_13` column gives the
temperature at 1pm. Note that both `pollution` and `wndspd` have been
$\log\left(1 + x\right)$ transformed.

    a. Cluster all the days using a hierarchical clustering. Be sure to scale
    the columns in advance, so that no single variable has undue influence on
    the result. Cut the tree so that we are left with 10 clusters, and provide a
    ribbon plot to visualize several quantiles associated with each cluster's
    typical pollution, temperature, and windspeed trajectories. visualization to
    understand the centroids of each cluster. Briefly interpret the main
    patterns you observe.
    
make the clustering
1. scale the data before clustering
2. Run hierarchical clustering
3. cut the tree to get 10 clusters
    
```{r}
scaled_pollution <- pollution %>%
  select(-date) %>%
  scale()
```

```{r}
fit <- scaled_pollution %>%
  dist() %>%
  hclust()
```

```{r}
memberships <- cutree(fit, 10)
 memberships
```
    
visualize the clustering (ribbon plot)
1. get quantiles associated with each cluster
2. visualize the quantiles

```{r}
pollution_ <- pollution %>%
  mutate(cluster = factor(memberships))
pollution_$cluster
```

```{r}
pollution_ <- pollution_ %>%
  pivot_longer(contains("_")) %>%
  separate(name, c("variable", "hour"), convert = TRUE)
```
    
```{r}
quantiles <- pollution_ %>%
  group_by(cluster, variable, hour) %>%
  summarise(
    q1 = quantile(value, 0.25),
    median = median(value),
    q3 = quantile(value, 0.75),
  )

quantiles
```
```{r}
ggplot(quantiles) +
  geom_ribbon(aes(hour, ymin = q1, ymax = q3), alpha = 0.3) +
  geom_line(aes(hour, median)) +
  facet_grid(variable ~ cluster, scale = "free_y")
```
    
```{r}
pollution <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week9/exercises/pollution_wide.csv")
pollution
```
    
    b. Visualize the hierarchical clustering tree associated with your
    clustering from part (a). Bonus: For a single time of day, encode the
    average pollution level for all descendant nodes within your tree
    representation. To summarize a the pollution level of all descendants leaves
    in a tree, you may use the `map_local_dbl` function from `tidygraph`.
    
```{r, eval = FALSE}
mean_fun <- function(neighborhood, cur_var = "pollution_10", ...) {
  cur_pollution <- neighborhood %>%
    filter(leaf) %>%
    pull(cur_var)
  mean(cur_pollution, na.rm = TRUE)
}

G %>%
  mutate(
    avg_pollution = map_local_dbl(order = graph_order(), mode = "out", .f = mean_fun)
    )
```