---
author: ""
title: "Week9W Code Starter"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

[Chicago Traffic Accidents] The dataset [here](https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week9/exercises/chicago_crashes.csv) is a preprocessed version of a
dataset downloaded from Chicago's official traffic accidents
[database](https://data.cityofchicago.org/Transportation/Traffic-Crashes-Crashes/85ca-t3if/data).
In this problem, we will use a heatmap to understand what times during the week
and over the year are associated with the highest number of traffic accidents.


    a. Use a heatmap to visualize the accident counts over time. Specifically,
    one dimension of the heatmap should index the 168 hours in the week, and the
    other should index every week from 2013-03-03 to 2022-03-27.
    
```{r}
library(tidyverse)
crashes <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week9/exercises/chicago_crashes.csv") %>%
  pivot_wider(names_from = c("day_of_week", "hour"), values_from = "n", values_fill = 0)
```

```{r}
crashes <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week9/exercises/chicago_crashes.csv")
crashes$hour %>% unique()
```




```{r, fig.height = 10}
library(superheat)
crashes_mat <- crashes %>%
  pivot_wider(names_from = c("day_of_week", "hour"), values_from = "n", values_fill = 0) %>%
  select(-week)
superheat(
  t(crashes_mat),
  pretty.order.cols = TRUE
)
```
    
    b. Enrich your visualization with information about the total number of
    accidents in each of the 168 hours in the week. _Hint: Use the `yt` or `yr`
    arguments to `superheat`._
  
```{r, fig.width = 6, fig.height = 4}
superheat(
  t(crashes_mat),
  pretty.order.cols = TRUE,
  yr = colSums(crashes_mat),
  yr.plot.type = "bar",
  yr.bar.col = "#d3d3d3",
  yr.obs.col = rep("#d3d3d3", ncol(crashes_mat))
)
```
  
  
    c. Provide a brief interpretion of your final visualization.
    