---
author: ""
title: "Week10M Code Starter"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```
This problem will use PCA to provide a low-dimensional view of a 14-dimensional
nutritional facts
[dataset](https://uwmadison.box.com/shared/static/nmgouzobq5367aex45pnbzgkhm7sur63.csv).
The data were originally curated by the USDA and are regularly used in
[visualization studies](http://bl.ocks.org/syntagmatic/raw/3150059/).

```{r}
nutrients <- read_csv("https://uwmadison.box.com/shared/static/nmgouzobq5367aex45pnbzgkhm7sur63.csv")
nutrients
```

a. Define a tidymodels `recipe` that normalizes all nutrient features and
specifies that PCA should be performed.

```{r}
library(tidymodels)
pca_recipe <- recipe(~., data = nutrients) %>%
 update_role(id:group_lumped, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors())

pca_prep <- prep(pca_recipe)
  # see bullet 10 here: https://krisrs1128.github.io/stat479/posts/2021-03-24-week10-2/
```

b. Visualize the top 6 principal components. What types of food do you expect to
have low or high values for PC1 or PC2?

```{r}
components <- tidy(pca_prep, 2) %>%
  filter(component %in% str_c("PC", 1:6)) %>%
  mutate(
    terms = str_remove(terms, " \\(g\\)")
  )
```


```{r, fig.height = 6, fig.width = 8}
library(tidytext)
ggplot(components) +
  geom_col(aes(value, reorder_within(terms, value, component))) +
  scale_y_reordered() +
  facet_wrap(~ component, scales = "free_y")
```

c. Compute the average value of PC2 within each category of the `group` column.
Give the names of the groups sorted by this average.

```{r}
scores <- juice(pca_prep)
group_order <- scores %>%
  group_by(group) %>%
  summarise(mean_pc = mean(PC2)) %>%
  arrange(mean_pc) %>%
  pull(group)

group_order
```

d. Visualize the scores of each food item with respect to the first two
principal components. Facet the visualization according to the `group` column,
and sort the facets according to the results of part (c). How does the result
compare with your guess from part (b)?

```{r, fig.width= 9, fig.height = 6}
scores <- scores %>%
  mutate(group = factor(group, levels = as.character(group_order)))

ggplot(scores) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_point(aes(PC1, PC2, col = group_lumped), size = 0.6, alpha = 0.5) +
  facet_wrap(~ group)
```
