---
author: ""
title: "Week10M Code Starter"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```
This problem will use PCA to provide a low-dimensional view of a 14-dimensional
nutritional facts
[dataset](https://uwmadison.box.com/shared/static/nmgouzobq5367aex45pnbzgkhm7sur63.csv).
The data were originally curated by the USDA and are regularly used in
[visualization studies](http://bl.ocks.org/syntagmatic/raw/3150059/).

```{r}
nutrients <- read_csv("https://uwmadison.box.com/shared/static/nmgouzobq5367aex45pnbzgkhm7sur63.csv")
```

a. Define a tidymodels `recipe` that normalizes all nutrient features and
specifies that PCA should be performed.

```{r}
pca_recipe <- recipe(~., data = nutrients) #%>%
  # see bullet 10 here: https://krisrs1128.github.io/stat479/posts/2021-03-24-week10-2/
```

b. Visualize the top 6 principal components. What types of food do you expect to
have low or high values for PC1 or PC2?

```{r}
tidy(pca_prep, 2)
```

c. Compute the average value of PC2 within each category of the `group` column.
Give the names of the groups sorted by this average.

d. Visualize the scores of each food item with respect to the first two
principal components. Facet the visualization according to the `group` column,
and sort the facets according to the results of part (c). How does the result
compare with your guess from part (b)?

```{r}
juica(pca_prep)
```
