---
author: ""
title: "Exercise 11.1 Code Starter"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```

Notes:

  * For this problem, it's enough to work with a sample of 500 documents.
  * Some functions that may be helpful are: `unnest_tokens()` and `cast_dtm()`.
  * I've included solutions for parts (c) and (d) in case you finish (a-b) early
  and want to try out the last two parts. I assume the fitted LDA object is in
  an object called `fit`.

This
[dataset](https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week11/exercises/mbti_1.csv?raw=true)
contains a sample of posts from the Personality Cafe forum, together with labels
giving the poster's (self-reported) Meyer-Briggs type (e.g., types starting with
"I" are introverted, those with "E" are extroverted). It is time-consuming to go
through all the posts manually, but to get a quick overview of the main types
that are discussed, we can use a Latent Dirichlet Allocation (LDA) model. This
problem walks through the steps for (1) preparing a Document-Term matrix for
LDA, (2) fitting the LDA model, and (3) interpreting the estimated topics and
memberships.

a. Transform the raw posts into a collection of per-post word counts. Remove
stopwords from across all lexicons in `tidytext::stopwords`.

For this exercise, it's enough to work with a sample of 500 documents. You may find the `unnest_tokens()` and `anti_join()` functions helpful

```{r}
library(tidytext)
mbti <- read_csv("https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week11/exercises/mbti_1.csv?raw=true") %>%
  sample_n(500) %>% # can set sample_n(50) for quick testing
  mutate(document = as.character(row_number()))

mbti
```

b. Convert the `data.frame` from (a) into a topic models Document-Term
matrix (i.e., an object of class `DocumentTermMatrix`). Fit an LDA model
with 8 topics to the prepared object.

c. Visualize the top 30 words within each of the fitted topics.
Specifically, create a faceted bar chart where the lengths of the bars
correspond to word probabilities and the facets correspond to topics.
Reorder the bars so that each topic's top words are displayed in order of
decreasing probability.

We first find the top 30 words per topic.

```{r}
beta <- tidy(fit, matrix = "beta") %>%
  group_by(topic) %>%
  slice_max(beta, n = 30) %>%
  mutate(term = reorder_within(term, beta, topic))
```

We can next visualize the words in each topic. It seems like many similar words
appear across all topics -- it might be more worthwhile to look for
discriminative words, as discussed in the case study lecture.

```{r, fig.height = 8, fig.width = 4}
ggplot(beta) +
  geom_col(aes(beta, term)) +
  scale_y_reordered() +
  facet_wrap(~ topic, scales = "free_y", ncol = 3)
```

d. Create a Structure plot displaying the topic memberships for each
document. Sort documents according to their order on a hierarchical
clustering tree, and facet documents according to personality type. Are
there certain topics that appear to be more common in some personality types
than others?

We first join in the personality type information with each document's membership.

```{r}
personality_types <- mbti %>%
  select(document, type) %>%
  unique()

memberships <- tidy(fit, matrix = "gamma") %>%
  left_join(personality_types)
```

We now reorder documents according to their similarity in topic memberships.

```{r}
hclust_result <- hclust(dist(fit@gamma))
document_order <- fit@documents[hclust_result$order]
memberships <- memberships %>%
  mutate(document = factor(document, levels = document_order))
```

We can finally create a structure plot.

```{r, fig.height = 10, fig.width = 6}
ggplot(memberships) +
  geom_col(aes(gamma, document, fill = as.factor(topic), col = as.factor(topic))) +
  facet_grid(type ~ ., scale = "free", space = "free") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_color_brewer(palette = "Set2") +
  scale_fill_brewer(palette = "Set2") +
  theme(
    strip.text.y = element_text(angle = 0),
    axis.text.y = element_blank(),
    axis.ticks = element_blank()
  )
```
