---
author: ""
title: "Week 11M"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
```


[Topics in _Pride and Prejudice_] This problem uses LDA to analyze the full text
of _Pride and Prejudice_. The object `paragraph` is a data.frame whose rows are
paragraphs from the book. We've filtered very short paragraphs; e.g., from
dialogue. We're interested in how the topics appearing in the book vary from the
start to the end of the book, for example.

```{r}
library(tidytext)
paragraphs <- read_csv("https://uwmadison.box.com/shared/static/pz1lz301ufhbedzsj9iioee77r95xz4v.csv")
paragraphs
```
      
a. Create a Document-Term Matrix containing word counts from across the same
paragraphs. That is, the $i^{th}$ row of `dtm` should correspond to the $i^{th}$
row of `paragraph`. Make sure to remove all stopwords.

```{r}
stop_words
```


```{r}
library(tidytext)
paragraphs_dtm <- paragraphs %>%
  unnest_tokens(words, text) %>%
  filter(! (words %in% stop_words$word)) %>%
  count(paragraph, words) %>%
  cast_dtm(paragraph, words, n)

paragraphs_dtm
```


b. Fit an LDA model to `dtm` using 6 topics. Set the seed by using the argument
`control = list(seed = 479)` to remove any randomness in the result.

```{r}
library(topicmodels)
fit <- LDA(paragraphs_dtm, k = 6, control = list(seed = 479))

topics <- tidy(fit, matrix = "beta") %>%
  group_by(topic) %>%
  slice_max(beta, n = 30)

head(topics)
```


```{r, fig.height = 6, fig.width = 8}
topics %>%
  mutate(term = reorder_within(term, beta, topic)) %>%
  ggplot() +
  geom_col(aes(beta, term)) +
  scale_y_reordered() +
  facet_wrap(~ topic, scale = "free")
```

c. Visualize the top 30 words within each of the fitted topics.
Specifically, create a faceted bar chart where the lengths of the bars
correspond to word probabilities and the facets correspond to topics.
Reorder the bars so that each topic's top words are displayed in order of
decreasing probability.

d. Find the paragraph that is the purest representative of Topic 2. That is,
if $\gamma_{ik}$ denotes the weight of topic $k$ in paragraph $i$, then
print out paragraph $i^{\ast}$ where $i^{\ast} = \arg \max_{i}\gamma_{i2}$.
Verify that the at least a few of the words with high probability for this
topic appear. Only copy the first sentence into your solution.

