---
author: "Kris Sankaran"
title: "Week9M"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, dpi = 200, fig.align = "center", out.width = 650, fig.height = 3, fig.width = 9)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Visualizing $K$-means

.pull-left[
  March 28, 2022
]
 
---

### Announcements

* Portfolio 3 - April 10
* Project Milestone 3 - April 10

---

### Today

By the end of the class, you should be able to...

   * Apply pivoting to provide input data for the `kmeans()` function
   * Visualize the cluster centroids and memberships generated by `kmeans()`

---

### Exercise Review

In this problem, we'll study a network
dataset of Amazon bestselling US Politics books. Books are linked by an edge if
they appeared together in the recommendations ("customers who bought this book
also bought these other books").
    
---
    
a. The code below reads in the edges and nodes associated with the network.
The edges dataset only contains IDs of co-recommended books, while the nodes
data includes attributes associated with each book. Build a `tbl_graph`
object to store the graph.

```{r}
library(tidygraph)
edges <- read_csv("https://uwmadison.box.com/shared/static/54i59bfc5jhymnn3hsw8fyolujesalut.csv", col_types = "cci")
nodes <- read_csv("https://uwmadison.box.com/shared/static/u2x392i79jycubo5rhzryxjsvd1jjrdy.csv", col_types = "ccc")
G <- tbl_graph(nodes, edges)
G
```

---

b. Use the result from part (a) to visualize the network as a node-link
diagram. Include the book's title in the node label, and shade in the node
according to political ideology.

```{r, fig.width = 10, fig.height = 6, out.width = 800, echo = FALSE, fig.retina = 2}
library(ggraph)
ggraph(G, "kk") +
  geom_edge_link(width = 1.5, color = "#d3d3d3", alpha = 0.5) +
  geom_node_label(aes(fill = political_ideology, label = Label)) +
  scale_fill_manual(values = c("#F26849", "#5377A6", "#BFBFBF")) +
  theme_void()
```


b. Use the result from part (a) to visualize the network as a node-link
diagram. Include the book's title in the node label, and shade in the node
according to political ideology.

```{r, fig.width = 10, fig.height = 6, eval = FALSE}
library(ggraph)
ggraph(G, "kk") +
  geom_edge_link(width = 1.5, color = "#d3d3d3", alpha = 0.5) +
  geom_node_label(aes(fill = political_ideology, label = Label)) +
  scale_fill_manual(values = c("#F26849", "#5377A6", "#BFBFBF")) +
  theme_void()
```


---

c. Create the analogous adjacency matrix visualization. Provide examples of
visual queries that are easy to answer using one encoding but not the other
(i.e., what is easy to see in the node-link view vs. what is easy to see in
the adjacency matrix).

```{r, fig.width = 4, fig.height = 4.3, out.width = 400, echo = FALSE}
ggraph(G, "matrix") +
  geom_edge_tile(mirror = TRUE) +
  geom_node_point(aes(col = political_ideology), x = -1) +
  geom_node_point(aes(col = political_ideology), y = nrow(data.frame(G)) + 1) +
  scale_color_manual(values = c("#F26849", "#5377A6", "#BFBFBF")) +
  coord_fixed() +
  theme_void() +
  theme(legend.position = "bottom")
```

---

c. Create the analogous adjacency matrix visualization. Provide examples of
visual queries that are easy to answer using one encoding but not the other
(i.e., what is easy to see in the node-link view vs. what is easy to see in
the adjacency matrix).

```{r, fig.width = 4, fig.height = 5, eval = FALSE}
ggraph(G, "matrix") +
  geom_edge_tile(mirror = TRUE) +
  geom_node_point(aes(col = political_ideology), x = -1) +
  geom_node_point(aes(col = political_ideology), y = nrow(data.frame(G)) + 1) +
  scale_color_manual(values = c("#F26849", "#5377A6", "#BFBFBF")) +
  coord_fixed() +
  theme_void() +
  theme(legend.position = "bottom")
```

---

* Node-link clearly shows neutral books between conservative and liberal
* Node-link makes topological queries simple
* Adjacency matrix reveals tighter sub-clusters
* Adjacency matrix would be more useful if text could appear on mouseover

---

## Live Coding Example

We will work through Exercise 3 in Module 3.

---

### Exercise

* Exercise 9.1 [Colony Collapse] on Canvas
* Can discuss, but submit individually