---
author: "Kris Sankaran"
title: "Week9M"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, dpi = 200, fig.align = "center", out.width = 650, fig.height = 3, fig.width = 9)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Visualizing Hierarchical Clustering

.pull-left[
  March 30, 2022
]
 
---

### Today

By the end of the class, you should be able to...

   * Interpret the centroids and cluster memberships derived by $K$-means
   * Construct a visualization of a hierarchical clustering output

---

### Exercise Review
   
In this problem, we will study a dataset describing factors that might be
leading to colony collapse disorder among bees. Since there are multiple
stressors for each state $\times$ timepoint combination, we will use clustering
to create a summarized "ecological stress" profile.

---

  a. The code below reads in data from its original Tidy Tuesday source and
  replaces NAs with 0's. We will summarize each state $\times$ timepoint
  combination by its profile of `stress_pct` across each stressor. Reshape the
  data so that `stressor` appears along columns, with the associated
  percentage contained within the table. Each row is therefore a profile of
  the amount of different stressors at a particular location and time.
  
```{r}
library(tidyverse)
stressor <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-11/stressor.csv') %>%
  mutate(stress_pct = replace_na(stress_pct, 0))

stressor <- stressor %>%
  pivot_wider(names_from = "stressor", values_from = "stress_pct")
stressor
```
    
---
    
b. Apply $K$-means to the profiles constructed in part (a). You may choose
$K$. Visualize and briefly interpret the resulting centroids.

```{r}
fit <- stressor %>%
  select(`Varroa mites`:Unknown) %>%
  kmeans(centers = 3, nstart = 10)

fit
```

---

We first reshape the centroids so we can visualize them.

```{r}
centroids <- data.frame(fit$centers) %>%
  mutate(cluster = as.factor(row_number())) %>%
  pivot_longer(-cluster, "variable")

centroids
```

---

Each centroid (panel) is the average of samples in each cluster.

```{r fig.height = 3.8, fig.width = 8, out.width = 900}
ggplot(centroids) +
  geom_point(aes(value, reorder(variable, value))) +
  facet_wrap(~ reorder(cluster, value))
```

---

```{r, fig.height = 4, fig.width = 10, out.width = 900}
ggplot(centroids) +
  geom_tile(aes(variable, reorder(cluster, value), fill = value)) +
  scale_fill_distiller(direction = 1) +
  scale_x_discrete(expand = c(0, 0)) + 
  scale_y_discrete(expand = c(0, 0))
```

---

c. Design and implement a visualization to show the change in cluster
memberships over time. Briefly interpret your results in context.
        
```{r}
library(tidymodels)
stressor <- augment(fit, stressor) %>%
  pivot_longer(`Varroa mites`:Unknown)
stressor
```

---

```{r, fig.height = 4.4, fig.width = 8, out.width = 800}
stressor %>%
  count(year, months, .cluster) %>%
  ggplot() +
  geom_line(aes(year, n, col = .cluster)) +
  facet_wrap(~ months)
```

---

* Clusters counts are much easier to interpret
* But, we would not have noticed the all-missing April 2019

```{r, fig.height = 3.8, fig.width = 8, out.width = 800}
ggplot(stressor, aes(year, value, group = state)) +
    geom_line(col = "#0c0c0c", alpha = 0.6) +
    geom_point(aes(col = .cluster)) +
    facet_grid(name ~ months, scale = "free_y")
```
  
---

## Live Coding Example

We will work through Exercise 10 in Module 3.

---

### Exercise

* Exercise 9.2 [Interpreting Clustering] on Canvas
* Can discuss, but submit individually