---
author: "Kris Sankaran"
title: "week3F"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, dpi = 200, fig.align = "center", out.width = 650)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Compound Figures Review

.pull-left[
February 11, 2022
]
 
---

### Announcements

* Project Milestone 1 due Sunday evening
 
---

## Exercise 3.2 Discussion

---

### Community Knowledge

1. Think about the community you will engage with

  - What is the present body of knowledge?
  - How do people make decisions? Will you work influence it?
  - Avoid statements that are only about the data (and not about the context)
  
1. It will be important to read about your problem area

  - It will be easier to develop good questions if you are familiar with the
  topic
  - The most interesting ideas often just link existing work in surprising ways
  
---

### Imitation

1. Find work that inspires you, and try to deconstruct it

1. Don't be afraid of imitation, especially when learning a new area

  - By imitating someone's solution technique in your own problem, you can learn
  it more deeply
  - My first visualization [project](https://krisrs1128.github.io/OpioidAtlas)
  was a (poor) imitation of [work](http://nbremer.github.io/urbanization/) that
  I thought was beautiful

---

## Muddiest Points

---

* When is a dataset tidy?
* When we receive a dataset, how should we re-arrange the rows and columns?

---

1. Often, "tidyness" depends on the purpose of the analysis. 
1. Each mark on the page should correspond to a single row in the input
`data.frame`, but this may vary from figure to figure.
  - In a survey, we can tidy so that each survey respondent is in a row
  and questions are along columns (so $n$ rows, one per respondent).
  - Alternatively, we can reshape it so that each question response gets its own
  row (so $n \times Q$ rows, one per respondent-question pair).
  
---

```{r, fig.width = 6, fig.height = 4}
library(tidyverse)
plants <- read_csv("https://uwmadison.box.com/shared/static/qg9gwk2ldjdtcmmmiropcunf34ddonya.csv")
ggplot(plants, aes(height.7, height.21)) +
  stat_smooth(method = "lm", se = F) +
  geom_point()
```

---

```{r, fig.width = 6, fig.height = 2.5}
plants %>%
  pivot_longer(
    starts_with("height"), names_sep = "\\.", 
    names_to = c("prefix", "time"), 
    names_transform = list(time = as.integer) ,
    values_to = "height"
  ) %>%
  ggplot() +
    geom_line(aes(time, height, group = interaction(plantid, treatment), col = treatment))
```

---

1. In practice, create one folder with the raw data. Never touch this folder.
1. Prepare a script to generate a few versions of the "derived" data. These can
be tidied into several forms.
  - One derived dataset with respondents on rows
  - Another with question responses on rows
1. Downstream visualizations can refer to either derived dataset

---

### Notes review

(go to [link](https://drive.google.com/file/d/1r1HEhAn9SCxqicuX2VNbdUv3w-qQBeRj/view?usp=sharing))

---

## Exercise

---

### Options

* Ridgelines practice: German Traffic [Module 1, Problem 10]
* Compound figures practice: Weather [Module 1, Problem 25]

---

### Hints

* For the traffic dataset, consider using the amount of traffic (`value`) for the
`height` and the city (`name`) for the `y` argument in `geom_ridgeline()`
* It may be interesting to reorder the cities by some property of their traffic patterns
* For the weather dataset, make sure to distinguish between the two locations
* For the weather dataset, a `geom_ribbon` might be interesting to compare
`temp_min` and `temp_max`

---

### Exercise

* Exercise 3.3 on Canvas
* Discuss in groups, but submit own solution
* Until: 

