---
author: "Kris Sankaran"
title: "week3M"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, dpi = 200, fig.align = "center", out.width = 650)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Faceting Review

.pull-left[
February 7, 2022
]

---

### Announcements

* View exercises as opportunity for experimentation
* We will go over additional muddy points on Wednesday and Friday
* I'll start assigning project-related in-class exercises this week
 
---

## Exercise Review
 
---

### Part (a)

Reading in the data and addressing merged sports issue,

```{r}
olympics <- read_csv("https://uwmadison.box.com/shared/static/rzw8h2x6dp5693gdbpgxaf2koqijo12l.csv") %>%
  separate(Sport, c("Sport", "Sport2", "Sport3"), sep = ",")
```


---

### Part (a)

Using separate on its own splits on characters.

.pull-left[
```{r}
olympics %>%
  select(`Place of birth`)
```
]

.pull-right[
```{r}
olympics %>%
  separate(`Place of birth`, c("birth_city", "birth_country")) %>%
  select(starts_with("birth"))
```
]

---

We need to customize the separate argument. Here, we separate by `(` or `)`.

```{r}
olympics %>%
  separate(`Place of birth`, c("birth_city", "birth_country"), sep = "\\(") %>%
  separate(birth_country, c("birth_country", "tmp"), sep = "\\)") %>%
  select(starts_with("birth"))
```

---

### Part (b)

```{r}
olympics %>%
  group_by(Sport) %>%
  summarise(sd = sd(Age)) %>%
  arrange(-sd)
```

---

### Part (c)

One approach is to use reorder within `aes()`.

```{r, fig.height = 10, fig.width = 12, echo = TRUE, eval = FALSE}
ggplot(olympics) +
  geom_point(
    aes(Age, reorder(Sport, Age, sd)),
    position = position_jitter(height = 0.25),
    alpha = 0.7, size = 0.8
  )
```

---

```{r, fig.height = 8, fig.width = 12, echo = FALSE}
ggplot(olympics) +
  geom_point(
    aes(Age, reorder(Sport, Age, sd)),
    position = position_jitter(height = 0.25),
    alpha = 0.7, size = 0.8
  )
```


---

It is reasonable to use a boxplot to summarize the sports-level distributions.
It does reduce the amount of information displayed, but the main takeaways are
the same.

```{r, fig.height = 8, fig.width = 12, eval = FALSE}
ggplot(olympics) +
  geom_violin(aes(Age, reorder(Sport, Age, sd)))
```

---

```{r, echo = FALSE, fig.height = 8, fig.width = 12}
ggplot(olympics) +
  geom_boxplot(aes(Age, reorder(Sport, Age, sd)))
```

---

Instead of calling `reorder()`, it is also possible to code the `Sports`
variable as a factor and set its levels according to part (a).

```{r}
sports_order <- olympics %>%
  group_by(Sport) %>%
  summarise(sd = sd(Age)) %>%
  arrange(sd) %>%
  pull(Sport)

sports_order
```

---

Instead of calling `reorder()`, it is also possible to code the `Sports`
variable as a factor and set its levels according to part (a).

```{r, fig.height = 8, fig.width = 12, eval = FALSE}
olympics <- olympics %>%
  mutate(Sport = factor(Sport, levels = sports_order))

ggplot(olympics) +
  geom_boxplot(aes(Age, Sport))
```

---

```{r, fig.height = 8, fig.width = 12, echo = FALSE}
olympics <- olympics %>%
  mutate(Sport = factor(Sport, levels = sports_order))

ggplot(olympics) +
  geom_boxplot(aes(Age, Sport))
```

---

### Exercise 2.3

* Many creative encodings,
  - Reordering according to win frequency
  - Point color instead of transparency
  - Tile color, and discard size
  - Line plots for top teams
  - Bar plots ordered by win frequency
* Many clarity improvements
  - Labeling axes
  - Resizing to legibility

---

```{r}
win_props <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/exercises/data/understat_per_game.csv") %>%
  group_by(team, year) %>%
  summarise(n_games = n(), wins = sum(wins) / n_games)

best_teams <- win_props %>%
  ungroup() %>%
  slice_max(wins, prop = 0.2) %>%
  pull(team)
```

---

```{r, fig.width = 10, fig.height = 9, out.width = 500}
win_props <- filter(win_props, team %in% best_teams)
ggplot(win_props) +
  geom_point(aes(year, reorder(team, wins), size = n_games, col = wins)) +
  scale_color_gradient2(midpoint = 0.5)
```

---

```{r, fig.width = 10, fig.height = 9, out.width = 500}
win_props <- filter(win_props, team %in% best_teams)
ggplot(win_props) +
  geom_tile(aes(year, reorder(team, wins), fill = wins, col = wins)) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  scale_fill_gradient2(midpoint = 0.5) +
  scale_color_gradient2(midpoint = 0.5)
```

---

```{r, fig.width = 10, fig.height = 5}
library(ggrepel)
ggplot(win_props, aes(n_games, wins)) +
  geom_point() +
  geom_text_repel(aes(label = team), size = 4) +
  ylim(0.7, 0.85) +
  facet_wrap(~ year)
```

---

```{r, fig.height = 6, fig.width = 10}
ggplot(win_props) +
  geom_point(aes(wins, reorder(team, wins), size = n_games, col = year)) +
  scale_size(range = c(0, 2)) +
  theme(axis.text = element_text(size = 8))
```

---

### Notes Summary

See [link](https://drive.google.com/file/d/1uf_VN9sCKf5FibGLhpPneix_HlXZ2a7G/view?usp=sharing).

---

### Exercise [Pokemon]

This problem gives practice in deriving new variables to improve a faceted plot.
The data below give attack and defense statistics for Pokemon, along with their
types. We will build a visualization to answer the question – how do the
different types of Pokemon vary in their attack and defense potential?

a. Derive a new column containing the attack-to-defense ratio, defined as $\frac{\text{Attack}}{\text{Defense}}$.

c. Provide a plot of the attack vs. defense scores for each Pokemon, faceted by
`type_1`.

Bonus 1: Label those pokemon that have exceptionally high (or low) ratios.
Bonus 2: Sort the facets according to median attack to defense ratio.

```{r}
pokemon <- read_csv("https://uwmadison.box.com/shared/static/hf5cmx3ew3ch0v6t0c2x56838er1lt2c.csv")
```

---

### Exercise Hints

* An example for this problem is in the Canvas course image
* For Bonus 1: Remember the derivation methods we discussed last week.
* For Bonus 2: Faceting order can be controlled by the levels of the factor
defining the facets. See today's notes / recording for an example.

---

### Exercise

* Exercise 3.1 on Canvas
* Discuss in groups, but submit own solution
* Until 2:00pm, then regroup and answer questions
