---
author: "Kris Sankaran"
title: "Week8M"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
library(shiny)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, dpi = 200, fig.align = "center", out.width = 650, fig.height = 4, fig.width = 6)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Networks and Trees (Part 2)

.pull-left[
  March 23, 2022
]
 
---

### Today

By the end of the class, you should be able to...

  * Use the `ggraph` package to customize a network visualization
  * Use the `tidygraph` package to generate and visualize clusters in a network

---

### Exercise Review

How do movie genres relate to one another? Is romance + comedy a more common
combination than fantasy + adventure? We will use the dataset
[here](https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week8/exercises/data/movie_links.csv)
to answer these questions, using a node-link diagram inspired by (but much
simpler than) the [film
flowers](https://observablehq.com/@sxywu/introduction-to-svg-and-d3-js) project
by Shirley Wu.

---

Build a `tbl_graph` containing the movies and their links.

```{r}
library(tidygraph)
edges <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week8/exercises/data/movie_links.csv")
G <- tbl_graph(edges = edges)
G
```

---

b. Create a node-link diagram to represent the connections between movies and their genres.

```{r, fig.width = 10, fig.height = 5}
library(ggraph)
ggraph(G) +
  geom_edge_link() +
  geom_node_text(aes(label = name))
```

---

b. Create a node-link diagram to represent the connections between movies and their genres.

```{r, fig.width = 11, fig.height = 6, out.width = 690}
ggraph(G, "kk") +
  geom_edge_link() +
  geom_node_text(aes(label = name)) +
  theme_void()
```

---

b. Create a node-link diagram to represent the connections between movies and their genres.

```{r, fig.width = 11, fig.height = 6, out.width = 690}
ggraph(G, "kk") +
  geom_edge_link(colour = "#d3d3d3", width = 0.5, alpha = 0.6) +
  geom_node_text(aes(label = name)) +
  theme_void()
```

---

```{r}
genres <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week8/exercises/data/genres.csv")
G <- G %>%
  activate(nodes) %>%
  mutate(type = name %in% genres$genre)
G
```

---

```{r, fig.width = 11, fig.height = 6, out.width = 730}
ggraph(G, "kk") +
  geom_edge_link(colour = "#d3d3d3", width = 0.5, alpha = 0.6) +
  geom_node_text(aes(label = name, size = type, col = type)) +
  scale_color_manual(values = c("#BF4545", "#225C73")) +
  scale_size_discrete(range = c(2.5, 5)) +
  theme_void() +
  theme(legend.position = "none")
```

---

## Live Coding Example

We will work through Exercise 20 in Module 2.

First, take 2 minutes to discuss one of the questions below,

  * What is your intuition about how the overall stack overflow tag network is
  structured? Do you expect to see hubs and spokes? Tightly connected clusters?
  Long trees or chains?
  * Imagine you were an engineer at Stack Overflow. What is one potential application of the tag network?
  * Imagine you are an academic studying how software engineers debug programs.
  What is one potential application of the tag network?
 
---

### Exercise

* Exercise 8.2 [Coltrane Network] on Canvas
* Can discuss, but submit individually