---
author: "Kris Sankaran"
title: "Week8M"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
library(shiny)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = FALSE, dpi = 200, fig.align = "center", out.width = 650, fig.height = 3, fig.width = 9)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Networks and Trees (Part 1)

.pull-left[
  March 21, 2022
]
 
---

### Announcements

* Portfolio 3 Peer Review - March 28
* Portfolio 3 - April 10
* Project Milestone 3 - April 10

---

### Today

By the end of the class, you should be able to...

  * Use the `ggraph` package to generate a network visualization
  * Use the `tidygraph` package to read in and manipulate a network dataset

---

### Common Issue in Portfolios

Q: How to use a brush along with other input widgets?

---

### Common Issue in Portfolios

Q: How to use a brush along with other input widgets?

* Be deliberate about how the user inputs should interact with one another.
Should one over-rule the other? Do you want to show the intersection or union of
rows satisfying the queries?
* Assuming we want an intersection or union, the simplest strategy is to observe
changes to any inputs, and update reactive values accordingly.

---

This is from a modified version of the Airbnb brushing example,

```{r, eval = FALSE}
server <- function(input, output) {
  selected <- reactiveVal(rep(TRUE, nrow(rentals)))
  observe({
    current <- update_selection(rentals, input$plot_brush, input$room_type)
    selected(current)
  })
  
  output$histogram <- renderPlot(overlay_histogram(rentals, selected()))
  output$map <- renderPlot(scatterplot(rentals, selected()))
}
```

---

This is from a modified version of the Airbnb brushing example,

```{r}
update_selection <- function(df, plot_brush, room_input) {
  sel1 <- brushedPoints(df, plot_brush, allRows = TRUE)$selected_
  sel2 <- df %>%
    mutate(selected_room = room_type %in% room_input) %>%
    pull(selected_room)
  
  sel1 & sel2
}
```

---

## Live Coding Example

We will work through Exercise 17 in Module 2.

First, take 2 minutes to discuss one of the questions below,

 * What is your current understanding of COVID-19's evolutionary history? What
 is one gap of your knowledge that this dataset might help fill?
 * What are some limitations of the available phylogenetic data? What attributes
 would you have liked to have been collected? Should our visualization come with
 any caveats?
 
---

### Exercise

* Exercise 8.1 [Movie Genres] on Canvas
* Can discuss, but submit individually