---
author: "Kris Sankaran"
title: "Week8F"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
library(shiny)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = FALSE, dpi = 200, fig.align = "center", out.width = 650, fig.height = 3, fig.width = 9)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Networks and Trees (Part 3)

.pull-left[
  March 25, 2022
]
 
---

### Today

By the end of the class, you should be able to...

  * Use the `ggraph` package to generate adjacency matrix visualizations
  * Discuss the relative merits of adjacency matrix vs. node-link visualizations

---

### Exercise Review

We will visualize the network of musicians who share a recording with John
Coltrane, obtained from this
[source](https://observablehq.com/@arnauqc/john-coltranes-co-authorship-network).
Two musicians are linked with one another if they appeared on an album together,
at least as recorded in the MusicBrainz database. To visualize these data, we
will practice creating network visualizations and encoding detected cluster
memberships.


---

a. Build a `tbl_graph` object based on the
[node](https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week8/exercises/data/coltraine-nodes.csv)
and
[link](https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week8/exercises/data/coltraine-edges.csv)
data.

```{r}
library(tidygraph)
nodes <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week8/exercises/data/coltraine-nodes.csv")
edges <- read_csv("https://raw.githubusercontent.com/krisrs1128/stat479_s22/main/_slides/week8/exercises/data/coltraine-edges.csv")
G <- tbl_graph(nodes, edges, directed = FALSE)
G
```

---

b. Create a node-link visualization associated with the network.

```{r, fig.width = 10, fig.height = 6}
library(ggraph)
ggraph(G, "fr") +
  geom_edge_link(colour = "#d3d3d3", width = 1, alpha = 0.8) +
  geom_node_label(aes(label = id), fill = "#f7f7f7") +
  theme_void()
```

---

c. Cluster nodes and encode modify your visualization from (b) to encode
cluster membership. _Hint: Mutate the graph using the `group_louvain()` or
`group_infomap()` functions from `tidygraph`._

```{r, fig.width = 10, fig.height = 6, out.width = 425}
G <- G %>%
  mutate(cluster = as.factor(group_louvain()))
ggraph(G) +
  geom_edge_link(colour = "#d3d3d3", width = 1, alpha = 0.8) +
  geom_node_label(aes(label = id, fill = cluster)) +
  scale_fill_brewer(palette = "Set2") +
  theme_void()
```


---

d. Comment on the visualization from (c). How would you interpret the resulting clustering?

* Two of the clusters seem relatively well-defined (blue and orange, though blue
has more within-cluster links).
* People who only collaborate with John Coltrane are assigned to the blue-green cluster.
* A bit surprisingly, Miles Davis, the Miles Davis Quintet, the Miles Davis
Sextet, and the Miles Davis All Stars do not all appear in the same cluster. The
same is true for Paul Chambers. But on some reflection, it makes sense -- two
groups including Miles Davis cannot possibly have had a collaboration together.

---

## Live Coding Example

We will work through (parts of) Exercise 18 in Module 2.

First, take 2 minutes to discuss the questions below,

  * What is your intuition about how this network is structured? If we had a few
  different sessions browsing videos, would we see several long chains? A few hubs and spokes? Tightly connected clusters?
  * What characteristics do you think result in links between videos? Similarity
  in topics? Being part of the same channels? The target having many more views?
 
---

### Exercise

* Exercise 8.3 [Political Book Recommendations] on Canvas
* Can discuss, but submit individually