---
author: "Kris Sankaran"
title: "Week13F"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
library(caret)
opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  cache = TRUE,
  dpi = 200,
  fig.align = "center",
  out.width = 650,
  fig.height = 3,
  fig.width = 9
  )
th <- theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Deep Learning Visualization

.pull-left[
  April 29, 2022
]

---

### Announcements

* Poster sessions on May 4, 6
* Portfolio 4 due May 8
* Project Milestone 4 due May 8

---

## Exercise Review

[Representation Analysis for CIFAR10] This problem asks you to investigate the
features learned by a deep learning model trained to the [CIFAR10
dataset](https://www.cs.toronto.edu/~kriz/cifar.html). We will study the
structure of the dataset and develop some intuition for what representations
different neurons are learning.
  
```{r, echo = TRUE, include = TRUE}
library(keras)
cifar <- dataset_cifar10()$train
f <- tempfile()
download.file("https://uwmadison.box.com/shared/static/c9kkxyrjb9myfj5knj5vnx0j8gn1uh0s.h5", f)
model <- load_model_hdf5(f)
```

---
  
c. The block below extracts feature activations associated with layer 7, for the
first 10% of the dataset. Specifically, the $ij^{th}$ element of `features`
gives the activation of neuron $j$ (in layer 7) on image $i$. Visualize the 10
images that have the highest activation for neuron 1. What does this neuron seem
to be responsive to?
    
```{r, echo = TRUE, include = TRUE}
activation_model <- keras_model(inputs = model$input, outputs = model$layers[[7]]$output)
features <- activation_model(cifar$x[1:5000,,, ]) %>%
  as.matrix()
```

---

```{r, fig.height = 8, fig.width = 12}
library(superheat)
log(1 + features) %>%
  t() %>%
  superheat(pretty.order.cols = TRUE, pretty.order.rows = TRUE)
```

---

* Strangely enough, this model picks up on military planes and small dogs
* It seems like the model picked up on several versions of the same image and
maybe even several angles of the same dog (?)

```{r, out.width = 100, fig.width = 10, fig.height = 10}
activation_order <- order(features[, 1], decreasing = TRUE)
cifar$x[activation_order[1:10],,,] %>%
  array_tree(1) %>%
  map(~ plot(as.raster(., max = 255)))
```

---

## Aside: Installation Errors

* The reason many of you encountered difficulties running `keras` from R is that
the `keras` package is actually running python code in the background
  - Advantage: Means that advances in python implementation are immediately accessible from R
  - Disadvantage: Linking package installation across both R and python is complex

---

## Live Coding Example

We will work through Exercise [Histopathology Embeddings] in Module 3.

---

### Exercise

Pick *one* example from [Model Visualization Use Cases] to discuss with your
project team. Submit your response as a group.

