---
author: "Kris Sankaran"
title: "Week13M"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
library(caret)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, dpi = 200, fig.align = "center", out.width = 650, fig.height = 3, fig.width = 9)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Deep Learning Visualization

.pull-left[
  April 25, 2022
]

---

## Exercise Review

[Overfitting `sin(x)`] We can use contrastive PD plots to visually evaluate
model over / underfitting. To illustrate, we will fit two $K$-nearest neighbors
models on a simple dataset, simulated below.
      
```{r, out.width = "0.4\\textwidth", fig.width = 6,, fig.height = 3, fig.align = "center", fig.cap = "A simulated sine wave dataset, used to evaluate overfitting."}
N <- 100
df <- data.frame(x = runif(N, -pi, pi)) %>%
  mutate(y = sin(x) + rnorm(N, 0, 0.2))

ggplot(df) +
  geom_point(aes(x, y))
```

---

Below, we train three models to fit this curve, averaging over the 2, 10, and
50 nearest neighbors before making a prediction at a new location.

```{r}
library(caret)

hyper <- map(c(2, 10, 50), ~ data.frame(k = .))
fits <- map(
  hyper, 
  ~ train(x = df %>% select(x), y = df$y, method = "knn", tuneGrid = .)
)

fits
```

---

a. Create an `explainer` object for each of the fits above. Use these to create
contrastive PD profiles between the three models.

```{r}
library(DALEX)
x <- df %>% select(starts_with("x"))
explanations <- imap(fits, ~ explain(.x, data = x, y = df$y, label = .y))
profiles <- map(explanations, model_profile)
plot(profiles[[1]], profiles[[2]], profiles[[3]])
```

---

b. Interpret the result from (a). Which models seem over / underfit? Which
would you use in practice, and how did you draw this conclusion?

The model with $K = 50$ is underfit -- it "fills the valleys" and "trims the
hills." The model with $K = 2$ seems to include the noise in the fitted
regression.

```{r}
plot(profiles[[1]], profiles[[2]], profiles[[3]])
```

---

c. Repeat the same analysis with a random forest model (`method = "rf"` in
caret). Instead of varying the `k` parameter, try several values of `mtry`. Is
this class of models more or less sensitive to overfitting than $K$-nearest
neighbors?
  
```{r}
hyper <- map(c(10, 50, 200), ~ data.frame(mtry = .))
fits <- map(
  hyper, 
  ~ train(x = df %>% select(x), y = df$y, tuneGrid = .)
)
```

---

The `mtry` parameter doesn't seem to have much influence in this problem.

```{r}
explanations <- imap(fits, ~ explain(.x, data = x, y = df$y, label = .y))
profiles <- map(explanations, model_profile)
plot(profiles[[1]], profiles[[2]], profiles[[3]])
```

---

## Live Coding Example

We will work through Exercise [Deep Learning as a Mixture of Separating Hyperplanes] in Module 3.

---

### Exercise

* Exercise 13.1 [Tensorflow Playground] on Canvas
* Can discuss, but submit individually
