---
author: "Kris Sankaran"
title: "Week7W"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
library(shiny)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = FALSE, dpi = 200, fig.align = "center", out.width = 650, fig.height = 3, fig.width = 9)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Geospatial Visualization (Part 2)

.pull-left[
  March 9, 2022
]
 
---

### Announcements

* Portfolio 2 - March 13
  - Separate pdf for discussion and zip file for app
  - Share both on piazza
  - Only unzipped directory needs to be runnable using `runApp`
* Project Milestone 2 - March 13

---

### Today

By the end of the class, you should be able to...

  * Read in and explain the summaries in a `raster` object
  * Create a polished, static visualization of a `raster` object using either
  `geom_raster` or `ggRGB`

---

### Exercise 7.1 Review

[Geospatial Commands] For each of the commands below, explain (i) whether it can
be applied to vector data, raster data, or both and (ii) discuss a typical use
case for the function.

a. `read_sf`

b. `raster`

c. `as.data.frame(img, xy = TRUE)`

d. `geom_sf`

e. `coord_sf`

---

a. `read_sf`

  (i) This applies to vector data only. 
  
  (ii) The command can be used to read vector data formats, like `.shp`, `.geojson`, into spatial `sf` objects.

[NYC Building Footprints](https://uwmadison.box.com/shared/static/qfmrp9srsoq0a7oj0e7xmgu5spojr33e.geojson)

```{r}
library(sf)
buildings <- read_sf("https://uwmadison.box.com/shared/static/qfmrp9srsoq0a7oj0e7xmgu5spojr33e.geojson")
str(buildings)
```

---

b. `raster`

(i) This applies to raster data only

(ii) This is the analogous reading function for raster data. The `extent` below
gives the latitude / longitude extent. `values` gives the population per pixel.

```{r}
library(raster)
pop <- raster("https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week7/exercises/data/afripop2020.tif?raw=true")
pop
```

---

c. `as.data.frame(img, xy = TRUE)`

(i) This applies to raster data only.

(ii) It is often used to convert a raster dataset to a data.frame for use in a
ggplot visualization.
```{r}
as.data.frame(pop, xy = TRUE) %>%
  drop_na() %>%
  head()
```

---

d. `geom_sf`

(i) This applies to vector data only

(ii) This is often used to add an `sf` layer to a ggplot visualization

```{r, out.width = 1200}
ggplot(buildings) +
  geom_sf(fill = "darkgray", size = 0)
```

---

e. `coord_sf`

(i) This applies to vector data only

(ii) This is often used to change the geographic limits of a ggplot
visualization or modify the coordinate reference system used

```{r}
ggplot(buildings) +
  geom_sf(fill = "darkgray", size = 0) +
  coord_sf(ylim = c(40.71, 40.72))

```

---

## Live Coding Example

We will work through Exercise 12 in Module 2


---

### Exercise

* Exercise 7.2 on Canvas
* Can discuss, but submit individually
* Hints:
  - Problem 5 requires the `sf` package
  - Problem 11 requires the `raster` package
  - For Problem 11, population density is highly skewed. Consider taking a
  $\log\left(1 + x\right)$ transformation of
  population