---
author: "Kris Sankaran"
title: "Week7M"
output:
  xaringan::moon_reader:
    css: ["default", "css/xaringan-themer.css"]
    self_contained: false
    lib_dir: libs
    nature:
      ratio: "16:9"
    seal: false
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
library(tidyverse)
library(shiny)
opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, dpi = 200, fig.align = "center", out.width = 650, fig.height = 3, fig.width = 9)
th <- theme_minimal() + 
  theme(
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "#f7f7f7"),
    panel.border = element_rect(fill = NA, color = "#0c0c0c", size = 0.6),
    axis.text = element_text(size = 14),
    axis.title = element_text(size = 16),
    legend.position = "bottom"
  )
theme_set(th)
options(width = 100)
```
class: bottom

# Geospatial Visualization (Part 3)

.pull-left[
  March 7, 2022
]
 
---

### Announcements

* Portfolio 2 - March 13
* Project Milestone 2 - March 13

---

### Today

By the end of the class, you should be able to...

  * Create a leaflet-based interactive map
  * Customize the styling for a leaflet-based map

---

### Exercise 7.2 Review - Option A
    
1. [Population Density] The problem below visualizes [population
data](https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week7/exercises/data/afripop2020.tif?raw=true)
from the `afrilearndata` package.

```{r}
library(raster)
population <- raster("https://github.com/krisrs1128/stat479_s22/blob/main/_slides/week7/exercises/data/afripop2020.tif?raw=true")
```

---

a. What are the dimensions (in pixels) of the population dataset? What is its
coordinate reference system?

Dimensions: 434 x 413 pixels
CRS: WGS84

```{r}
population
```

---
    
b. Convert the raster data to a `data.frame`. Omit rows without any
population data.

```{r}
population %>%
  as.data.frame(xy = TRUE)
```

---
    
b. Convert the raster data to a `data.frame`. Omit rows without any population
data.

```{r}
population %>%
  as.data.frame(xy = TRUE) %>%
  drop_na()
```

  
---
    
c. Visualize the data. What are some regions with the lowest and highest
population densities?
 

```{r, fig.width = 8, fig.height = 8, out.width = 290}
pop_df <- population %>%
  as.data.frame(xy = TRUE) %>%
  drop_na() %>%
  rename(population = afripop2020.tif.raw.true)
ggplot(pop_df) +
  geom_raster(aes(x, y, fill = population)) +
  coord_sf()
```

---
    
c. Visualize the data. What are some regions with the lowest and highest
population densities?
 

```{r, fig.width = 8, fig.height = 8, out.width = 390}
ggplot(pop_df) +
  geom_raster(aes(x, y, fill = log(1 + population))) +
  coord_sf()
```

---

```{r, fig.width = 8, fig.height = 8, out.width = 450}
ggplot(pop_df) +
  geom_raster(aes(x, y, fill = log(1 + population))) +
  coord_sf() +
  scale_fill_distiller(palette = 3, direction = 1) +
  theme_void()
```

---

.pull-left[
```{r}
ggplot(pop_df) +
  geom_histogram(aes(population))
```
]

.pull-right[
```{r}
ggplot(pop_df) +
  geom_histogram(aes(log(1 + population)))
```
]


---

### Exercise 7.2 Review - Option B

```{r, fig.width = 10, fig.height = 4, out.width = 900}
trees <- read_csv("https://uwmadison.box.com/shared/static/t1mk6i4u5ks5bjxaw2c7soe2z8i75m2o.csv")
ggplot(trees) +
  geom_point(aes(longitude, latitude, col = species_group)) +
  coord_fixed() +
  facet_wrap(~ health)
```


---

```{r, out.width = 1100, fig.height = 4, fig.width = 10, echo = FALSE}
trees <- trees %>%
  mutate(health = factor(health, levels = c("Good", "Fair", "Poor")))
ggplot(trees) +
  geom_point(
    aes(longitude, latitude, col = species_group, size = tree_dbh),
    alpha = 0.5
    ) +
  scale_color_brewer(palette = "Set3") +
  scale_size(range = c(0.01, 2)) +
  facet_wrap(~ health) +
  coord_fixed() +
  guides(col = guide_legend(override.aes = list(alpha = 1, size = 2), ncol = 3)) +
  theme(
    legend.position = "bottom",
    axis.text = element_text(size = 10)
  )
```

---

```{r, eval = FALSE}
trees <- trees %>%
  mutate(health = factor(health, levels = c("Good", "Fair", "Poor")))
ggplot(trees) +
  geom_point(
    aes(longitude, latitude, col = species_group, size = tree_dbh),
    alpha = 0.5
    ) +
  scale_color_brewer(palette = "Set3") +
  scale_size(range = c(0.01, 2)) +
  facet_wrap(~ health) +
  guides(col = guide_legend(override.aes = list(alpha = 1, size = 2), ncol = 3)) +
  theme(
    legend.position = "bottom",
    axis.text = element_text(size = 10)
  )
```

---

c. Generate a version of the plot in (a) that has the roads and buildings in the
    background.
    
```{r, out.width = 620, fig.width = 10, fig.height = 4}
library(sf)
roads <- read_sf("https://uwmadison.box.com/shared/static/28y5003s1d0w9nqjnk9xme2n86xazuuj.geojson")
buildings <- read_sf("https://uwmadison.box.com/shared/static/qfmrp9srsoq0a7oj0e7xmgu5spojr33e.geojson")

ggplot(trees) +
  geom_sf(data = roads) +
  geom_sf(data = buildings) +
  geom_point(aes(longitude, latitude, col = species_group)) +
  facet_wrap(~ health)
```

---

```{r, out.width = 1100, fig.height = 4, fig.width = 10, echo = FALSE}
ggplot(trees) +
  geom_sf(data = roads, size = .1) +
  geom_sf(data = buildings, size = 0) +
  geom_point(
    aes(longitude, latitude, col = species_group, size = tree_dbh),
    alpha = 0.5
  ) +
  scale_color_brewer(palette = "Set3") +
  scale_size(range = c(0.01, 2)) +
  scale_y_continuous(expand = c(0, 0), breaks = c(40.71, 40.725, 40.74)) +
  scale_x_continuous(expand = c(0, 0), breaks = c(-73.995, -73.975)) +
  facet_wrap(~ health) +
  guides(col = guide_legend(override.aes = list(alpha = 1, size = 2), ncol = 3)) +
  theme(
    legend.position = "bottom",
    axis.text = element_text(size = 10)
  )
```

---

```{r, eval = FALSE}
ggplot(trees) +
  geom_sf(data = roads, size = .1) +
  geom_sf(data = buildings, size = 0) +
  geom_point(
    aes(longitude, latitude, col = species_group, size = tree_dbh),
    alpha = 0.5
  ) +
  scale_color_brewer(palette = "Set3") +
  scale_size(range = c(0.01, 2)) +
  scale_y_continuous(expand = c(0, 0), breaks = c(40.71, 40.725, 40.74)) +
  scale_x_continuous(expand = c(0, 0), breaks = c(-73.995, -73.975)) +
  facet_wrap(~ health) +
  guides(col = guide_legend(override.aes = list(alpha = 1, size = 2), ncol = 3)) +
  theme(
    legend.position = "bottom",
    axis.text = element_text(size = 10)
  )
```

---

## Live Coding Example

We will work through Exercise 13 in Module 2.


---

### Exercise


---